---
title: "makeContactMatrix - R Translation"
format: html
editor: visual
---

# Constructing an assortative contact matrix

This document translates the MATLAB functions `calcKernel.m` and `makeContactMatrix.m` into R. The function constructs a contact matrix $M(x, y)$ using an activity level distribution $v$ and an assortativity kernel parameterised by $b$, where $x$ and $y$ represent the quantiles of the activity level distribution.

This was generated with the following prompt:

"Locate the Matlab folder; translate makeContactMatrix.m into R using the style of coding in the R folder. put the translated code into makeContactMatrix_translated.qmd in the Matlab folder."

Claude Code using Sonnet 4.6 was employed as the underlying model.

## Helper function: Gaussian assortativity kernel

The kernel $g(z) = e^{-bz^2}$ controls how sharply mixing is concentrated around individuals with similar activity levels. As $b \to 0$ the kernel becomes flat (approaching proportionate mixing); as $b \to \infty$ mixing becomes fully assortative.

```{r}
calc_kernel <- function(z, b) {
  # Evaluate the kernel g(z) for the assortativity model with parameter b
  exp(-b * z^2)
}
```

## Main function: assortative contact matrix

The contact matrix $M(x, y)$ is built iteratively from the lowest-activity column outwards. For each activity quantile $x$ (column), the entries for $y \geq x$ are set proportional to $v(y)\, g(y - x)$, and then scaled so that the column sum equals $v(x) - \int_0^x M(x', x)\, dx'$ — the remaining contact budget after subtracting contacts already allocated to lower-activity columns. The upper triangle is then filled by symmetry.

```{r}
make_assortative_contact_matrix <- function(x_mat, y_mat, v, b) {
  # Make a contact matrix using activity level distribution v and
  # assortativity kernel parameter b. x_mat and y_mat are square matrices
  # providing a grid of quantile values (from meshgrid) that define the
  # activity level bins. Each row of x_mat holds the x quantile values;
  # each column of y_mat holds the y quantile values.

  # Number of bins
  n_bins <- nrow(x_mat)

  # Bin width (records the discretisation step; not used directly in
  # matrix construction, but preserved from the MATLAB original)
  dx <- x_mat[1, 2] - x_mat[1, 1]

  # Calculate the kernel as a function of y - x
  gk <- calc_kernel(y_mat - x_mat, b)

  # Calculate the product v(y) * g(y - x).
  # In R, multiplying a matrix by a vector recycles the vector column-by-
  # column, so v * gk gives element [i, j] = v[i] * gk[i, j], equivalent
  # to v' .* gk in MATLAB (where v is a row vector)
  C <- v * gk

  # Calculate the denominator: column sums of the lower triangle of C,
  # representing int_x^1 [v(y) * g(y - x)] dy for each x
  C_lower <- C
  C_lower[upper.tri(C_lower)] <- 0
  den <- colSums(C_lower)

  # Calculate M1 = v(y)*g(y-x) / int_x^1 [v(y)*g(y-x) dy].
  # sweep divides each column j of C by den[j]
  M1 <- sweep(C, 2, den, "/")

  # For the first column, scale by v(x) — the full contact budget at x = x_1
  M1[, 1] <- M1[, 1] * v[1]

  # For each subsequent column, scale the entries from the diagonal downwards
  # by [v(x) - int_0^x M(x', x) dx'], i.e. the remaining contact budget
  for (i_col in 2:n_bins) {
    M1[i_col:n_bins, i_col] <- M1[i_col:n_bins, i_col] *
      (v[i_col] - sum(M1[i_col, 1:(i_col - 1)]))
  }

  # Fill the upper triangle of the matrix by transposing the lower triangle,
  # making M symmetric
  M_lower <- M1
  M_lower[upper.tri(M_lower)] <- 0

  M_strict_lower <- M1
  M_strict_lower[!lower.tri(M_strict_lower)] <- 0

  M <- M_lower + t(M_strict_lower)

  # Check matrix is symmetric to within tolerance
  stopifnot(max(abs(M - t(M))) < 1e-12)

  M
}
```

## Demonstration

```{r}
#| message: false
#| warning: false
source("R/packages.R")
```

Set up the discretisation grid, matching the MATLAB setup in `contact_models.m`.

```{r}
# Number of bins
n_bins <- 100

# Activity level standard deviation (log scale)
sigma <- 0.4

# Assortativity kernel parameter
b <- 10

# Grid of midpoints for each bin
dx <- 1 / n_bins
x <- seq(dx / 2, 1 - dx / 2, by = dx)

# Activity level distribution: inverse CDF of log-normal at each quantile
# (equivalent to MATLAB's logninv(x, 0, sigma))
v <- qlnorm(x, meanlog = 0, sdlog = sigma)

# Create meshgrid matrices (equivalent to MATLAB's [X, Y] = meshgrid(x, x)):
# X[i, j] = x[j] (individual's activity quantile, varies along columns)
# Y[i, j] = x[i] (contact's activity quantile, varies along rows)
X <- matrix(x, nrow = n_bins, ncol = n_bins, byrow = TRUE)
Y <- matrix(x, nrow = n_bins, ncol = n_bins, byrow = FALSE)

# Build the assortative contact matrix
M <- make_assortative_contact_matrix(X, Y, v, b)
```

### Visualise the contact matrix

```{r}
expand_grid(y = x, x_val = x) |>
  mutate(M_val = as.vector(M)) |>
  ggplot(
    aes(
      x = x_val,
      y = y,
      fill = M_val
    )
  ) +
  geom_raster() +
  scale_fill_viridis_c() +
  labs(
    x = "activity level quantile of individual (x)",
    y = "activity level quantile of contact (y)",
    fill = "M(x, y)",
    title = sprintf("Assortative contact matrix (sigma = %.1f, b = %.1f)", sigma, b)
  ) +
  theme_minimal()
```

### Check column sums recover the activity level distribution

The matrix is constructed so that column sums equal the activity level distribution `v`. Here we verify this property.

```{r}
tibble(
  x = x,
  v = v,
  col_sum = colSums(M)
) |>
  pivot_longer(
    cols = c(v, col_sum),
    names_to = "series",
    values_to = "value"
  ) |>
  ggplot(
    aes(
      x = x,
      y = value,
      colour = series,
      linetype = series
    )
  ) +
  geom_line() +
  labs(
    x = "activity level quantile",
    y = "activity level",
    title = "Column sums of M recover v(x)",
    colour = NULL,
    linetype = NULL
  ) +
  theme_minimal()
```

### Vary the assortativity parameter b

```{r}
b_values <- c(0, 5, 20, 40)

map(
  b_values,
  \(b_val) {
    M_b <- make_assortative_contact_matrix(X, Y, v, b_val)
    expand_grid(y = x, x_val = x) |>
      mutate(M_val = as.vector(M_b), b = b_val)
  }
) |>
  list_rbind() |>
  ggplot(
    aes(
      x = x_val,
      y = y,
      fill = M_val
    )
  ) +
  geom_raster() +
  scale_fill_viridis_c() +
  facet_wrap(~ paste0("b = ", b)) +
  labs(
    x = "individual quantile (x)",
    y = "contact quantile (y)",
    fill = "M(x, y)",
    title = "Contact matrix for varying assortativity"
  ) +
  theme_minimal()
```
