---
title: "Analysis S1 - Binning methods"
format: html
editor: visual
---

# Comparing alternative binning methods for summarising the activity level distribution

The computational efficiency of our proposed method relies on the ability to represent the population distribution of activity levels in a comparatively low number of bins. Specifically, we aim to rapidly converge on the true eigenvalue for a population, as the number of bins increases.

### Methods

This analysis compares a number of alternative methods:

#### Quantiles with mean activity
Define bins by evenly-spaced quantiles and approximate the mean activity levels in each bin

#### Gaussian quadrature scheme
Define bins according to the points of a Guassian quadrature scheme, with poulation fractions given by the normalised weights.

#### Evenly-sized bins
Defining bins of even size (i.e. range of activity levels between the upper and lower bound of the bin) up to the 99.9 percentile of the distribution, approximating the mean activity level (by numerical integration) and computing population fraction in each bin.

#### Hybrid scheme of quantiles and evenly-sized bins
Defining around half the bins to be equal quantiles, up to the 90 percentile, and then the remaining bins to be evenly sized, from the 90 to the 99.9 percentiles, approximating the mean activity level (by numerical integration) and computing population fraction in each bin.

### Evaluation

```{r}
#| message: false
#| warning: false
source("R/packages.R")
source("R/functions.R")
```

#### Approximating the distribution

First we assess the ability of the binning scheme to approximate the first two moments of the distribution, by computing numerical approximations to these parameters based on the binning schemes, and comparing them to the truth, for known values of sigma.

```{r}
estimates <- expand_grid(
  n_classes = 2:20,
  sigma = c(0.5, 1, 2),
  method = c("truth",
             "quantile_mean",
             "gaussquad",
             "quantile2_mean")
) |>
  rowwise() |>
  mutate(
    mean = case_when(
      method == "truth" ~ exp(0 + sigma ^ 2 / 2),
      .default = bin_estimate_mean(sigma,
                                   n_classes,
                                   method = method)
    ),
    sd = case_when(
      method == "truth" ~ sqrt((exp(sigma ^ 2) - 1) * exp(2 * 0 + sigma ^ 2)),
      .default = bin_estimate_sd(sigma,
                                 n_classes,
                                 method = method)
    )
  )

```

Plot these
```{r}
estimates |>
  pivot_longer(
    cols = c("mean", "sd"),
    values_to = "estimate",
    names_to = "moment"
  ) |>
  mutate(
    sigma = paste("sigma: ", sigma)
  ) |>
  ggplot(
    aes(
      x = n_classes,
      y = estimate,
      colour = method,
    )
  ) +
  geom_line() +
  facet_wrap(
    moment ~ sigma,
    scales = "free_y"
  ) +
  theme_minimal()

```
Perhaps unsurprisingly, Gaussian quadrature converges rapidly and consistently at this task. K-means appears to converge rapidly, but is quite stochastic (due to the random sampling used to fit the clustering model), with the other methods generally converging more slowly, especially for the standard deviation. The quantile median consistently understimates the standard deviation even with 20 bins.

#### Approximating the eigenvalue

While the above analysis is a fairly straightforward assessment of the ability of these schemes to integrate the distribution, the main task is to accurately estimate the eigenvalue of matrix constructed from this distribution. Given the strong influence of the upper tail of the activity distribution on the eigenvalue, the results of this analysis are likely to be quite different.

Approximate the eigenvalue for different numbers of bins for the different methods.

```{r}
plan(multisession, workers = 8)
sigma <- 1

eigenvalue_convergence <- expand_grid(
  bins = seq(2, 100, by = 1),
  method = c("quantile_mean",
             "gaussquad",
             "quantile2_mean")
) |>
  mutate(
    eigenvalue = furrr::future_map2(
      bins,
      method,
      ~ map_to_eigen(n_classes = .x,
                     sigma = sigma,
                     method = .y,
                     assort = 0),
      .options = furrr_options(seed = TRUE)
    )
  )
```

Plot convergence
```{r}
# this is the true eigenvalue for the proportional mixing assumption
true_eigenval <- exp(1.5 * sigma ^ 2)
eigenvalue_convergence |>
  unnest(
    eigenvalue
  ) |>
  ggplot(
    aes(
      x = bins,
      y = eigenvalue,
      group = method,
      colour = method)
  ) +
  geom_abline(
    intercept = true_eigenval,
    slope = 0,
    linetype = 2
  ) +
  geom_line() +
  xlab("Number of classes") +
  theme_minimal()
```

#### Monte Carlo "true" eigenvalue reference

To complement the analytical convergence curves above, we add a reference value computed directly from a large Monte Carlo sample (translation of `Matlab/matrix_bins.m`). We draw N = 10,000 individuals from a lognormal(0, sigma = 1) distribution, then compute the dominant eigenvalue of the full proportionate mixing matrix.

In MATLAB this required forming and diagonalising a 10,000 × 10,000 matrix (~800 MB). In R we exploit the rank-1 structure of the proportionate mixing matrix: for `M = a aᵀ / sum(a)`, the eigenvector is `a` itself and the dominant eigenvalue is `sum(a²) / sum(a)`. This closed form is exact and requires no matrix materialisation.

```{r}
set.seed(2024)
N <- 1e4

activity_mc <- rlnorm(N, meanlog = 0, sdlog = sigma)

# "True" MC eigenvalue: closed-form dominant eigenvalue of the rank-1
# proportionate mixing matrix M = a a^T / sum(a)
lambda_mc_true <- sum(activity_mc ^ 2) / sum(activity_mc)

# Population-level analytic value (the MC estimate converges to this as N -> Inf)
lambda_analytic <- exp(1.5 * sigma ^ 2)
```

Now compute the eigenvalue of binned approximations using quantile bins drawn from the MC sample, matching the MATLAB approach in `matrix_bins.m`.

```{r}
n_bins_range <- c(2:100, seq(200, 2000, by = 100))

eigenvalue_convergence_mc <- tibble(n_bins = n_bins_range) |>
  rowwise() |>
  mutate(
    eigenvalue = {
      probs <- seq(0, 1, length.out = n_bins + 1)
      bin_edges <- quantile(activity_mc, probs = probs)
      bin_edges[1] <- 0

      bin_means <- vapply(seq_len(n_bins), \(i) {
        in_bin <- activity_mc > bin_edges[i] & activity_mc <= bin_edges[i + 1]
        mean(activity_mc[in_bin])
      }, numeric(1))

      # Dominant eigenvalue of binned proportionate mixing matrix (rank-1 closed form)
      sum(bin_means ^ 2) / sum(bin_means)
    }
  ) |>
  ungroup()
```

Plot eigenvalue convergence with both MC and analytic reference lines.

```{r}
eigenvalue_convergence_mc |>
  ggplot(
    aes(x = n_bins, y = eigenvalue)
  ) +
  geom_line() +
  geom_point(size = 0.5) +
  geom_hline(
    yintercept = lambda_mc_true,
    linetype = "dashed",
    colour = "red"
  ) +
  geom_hline(
    yintercept = lambda_analytic,
    linetype = "dotted",
    colour = "blue"
  ) +
  annotate(
    "text", x = 200, y = lambda_mc_true * 1.004,
    label = '"true" MC value', colour = "red", hjust = 0
  ) +
  annotate(
    "text", x = 200, y = lambda_analytic * 0.997,
    label = "population analytic", colour = "blue", hjust = 0
  ) +
  scale_x_log10() +
  labs(
    x = "number of bins",
    y = "dominant eigenvalue",
    title = sprintf(
      "Eigenvalue convergence with MC-quantile binning (proportionate mixing, N = %d)", N
    )
  ) +
  theme_minimal()
```

The dashed red line is the finite-sample MC reference value; the dotted blue line is the population analytical value `exp(1.5 * sigma²)`. The binned eigenvalue converges to the MC reference, which itself fluctuates around the analytic value due to sampling variance from N = 10,000 draws.
