---
title: "Analysis 3 - Real-world application"
format: html
editor: visual
---

# Real-world analysis France and Hong Kong datasets

Computing and visualising the impact of the activity-heterogeneous analysis on analysis of datasets

-   Using conmat with an additional random effects term, estimate the activity-level distribution from HK & French data, and produce augmented (age and activity) contact matrices.

-   Using these, compute R0 under the augmented and non-augmented matrices (expect augmented to be higher)

-   Using these matrices, compute final size (using the `finalsize` package) for: the non-augmented matrix, the augmented matrix, and the non-augmented matrix rescaled to have the same R0 as the augmented matrix

-   Extend these analyses to do a bootstrap analysis to get uncertainty in these estimates

see Nick's conmat branch: `heterogeneous-augmentation` for partly complete implementation

NEED: Finish applying the heterogeneous-augmentation conmat analysis, and make figures

Load packages and functions

```{r}
#| message: false
#| warning: false
source("R/packages.R")
source("R/functions.R")

# set a seed for reproducibility, since contact ages are randomly imputed in
# process_fc_data_conmat()
set.seed(2026-07-06)
```

Install the experimental `heterogeneous-augmentation` branch of conmat for simultaneous fitting of age effects and residual heterogeneity.

```{r eval=FALSE}
remotes::install_github("goldingn/conmat@heterogeneous-augmentation")
```

Load French Connection data with the contact ages, and rename things to those expected by conmat

```{r}
fc_contact_data <- process_fc_data_conmat() |>
  rename(age_from = part_age,
         age_to = cont_age) |>
  mutate(
    setting = "all"
  )
```

Aggregate observations across all participants (this won't work for our social activity version, but doing this for the age-only conmat analysis)

```{r}
fc_contact_data_summarised <- fc_contact_data |>
  group_by(
    age_from,
    age_to,
    setting) |>
  summarise(
    contacts = sum(contacts),
    participants = n(),
    .groups = "drop"
  )
```

Get population data for France in 2010

```{r}
fc_population <- age_population(
  data = socialmixr::wpp_age(),
  location_col = country,
  location = c("France"),
  age_col = lower.age.limit,
  year_col = year,
  year = 2010
)
```

Run conmat on this and return the age-structured matrix, predicted to 5-year age groups up to 90, the oldest age among participants in the data.

```{r}
fc_model_aggregated <- fit_single_contact_model(
  contact_data = fc_contact_data_summarised,
  population = fc_population
)
max_age <- 90
age_breaks <- seq(0, max_age, by = 5)
age_contact_pred <- predict_contacts(fc_model_aggregated,
                                     fc_population,
                                     age_breaks = age_breaks)

age_matrix <- conmat::predictions_to_matrix(age_contact_pred)
age_matrix_plot <- autoplot(age_matrix)
ggsave("figures/fc_age_matrix.png",
       plot = age_matrix_plot,
       bg = "white",
       width = 6,
       height = 5,
       dpi = 300)
age_matrix_plot
```

This is the type of standard analysis we would use to estimate an age-structured contact matrix. Now we wish to estimate the amount of contact heterogeneity in the survey data that is not explained by this-structured contact matrix, parameterised by the residual standard deviation of (log-)contacts.

We have two methods for estimating sigma and the age effects:

1.  Two-stage inference for age structure and leftover between-individual variation:

    a\) estimate the age-structured contact matrix from the data, then

    b\) estimate the *residual* between-participant variation based on the predicted contact rates from step 1, using a random effects model.

2.  Use single model to simultaneously estimate age effects and inter-participant variation in contact rates.

Method 2 would be preferable since it correctly accounts for the residual of degrees of freedom in estimating both the age and activity effects, but method 1 does not require specialised software, can use any method for estimating an age structured contact matrix from the data, and is more computationally convenient. Attempt method 1 first.

Method 1 : a two-stage approach

We use conmat to estimate the age-structured contact matrix from these survey data. This age-only model (step a) has already been constructed above. For step b, we can fit a Poisson random effects model for the numbers of contacts in each age group for each participant, and include as an offset the predicted number of contacts from step a.\
\
First, obtain the predictions from conmat for the numbers of contacts for participants of any age to any other age:

```{r}
# range of ages we need to predict to
year_range <- range(c(fc_contact_data$age_from, fc_contact_data$age_to),
                    na.rm = TRUE)
# all the breaks
age_breaks_years <- seq(year_range[1], year_range[2] + 1, by = 1)

# make a lookup table to the age group notation
year_age_group_lookup <- tibble(
  age = age_breaks_years
) %>%
  mutate(
    age_group = sprintf("[%s,%s)", age, age + 1)
  )

# predict to these age groups, and add the integer age back on
age_contact_pred_years <- fc_model_aggregated %>%
  predict_contacts(
    fc_population,
    age_breaks = age_breaks_years
  ) %>%
  left_join(
    year_age_group_lookup,
    by = c(age_group_from = "age_group")
  ) %>%
  rename(
    age_from = age
  ) %>%
  left_join(
    year_age_group_lookup,
    by = c(age_group_to = "age_group")
  ) %>%
  rename(
    age_to = age,
    predicted_contacts = contacts
  ) %>%
  select(
    age_from,
    age_to,
    predicted_contacts
  )
```

Next, add this to the participant level data. We truncate participant and contact ages at `max_age` (as elsewhere), since the age-structured contact model extrapolates poorly in the sparsely-observed oldest age groups: predicted contact rates there collapse to near zero, so an observed contact against a near-zero offset would otherwise create extreme leverage that destabilises the random effects model.

```{r}
fc_contact_data_modelling <- fc_contact_data %>%
  left_join(
    age_contact_pred_years,
    by = c("age_from", "age_to")
  ) %>%
  filter(
    age_from <= max_age,
    age_to <= max_age
  )
```

Finally, fit a Poisson GLMM and extract the estimate of sigma.

```{r}
# fit the model. We use nAGQ = 0 (optimising the random effects within the
# PIRLS step rather than a separate adaptive Gauss-Hermite quadrature): the
# default nAGQ = 1 Laplace fit is numerically unstable here - the many rare
# age-pair observations with very small predicted offsets create high-leverage
# points that intermittently break the PIRLS step ("Downdated VtV is not
# positive definite"), depending on the random contact-age imputation. nAGQ = 0
# is robust to this and gives an effectively identical variance estimate.
model_twostage <- lme4::glmer(
  contacts ~ offset(log(predicted_contacts)) +
    (1|part_id),
  family = stats::poisson,
  data = fc_contact_data_modelling,
  nAGQ = 0
)
```

```{r}
# extract sigma
sigma_fc_twostage <- model_twostage |>
  summary() |>
  extracting("varcor") |>
  as_tibble() |>
  filter(grp == "part_id") |>
  select(sdcor) %>%
  pull()

sigma_fc_twostage
```

Method 2: simultaneous estimation

Now fit this with a single model, estimating age effects and activity level variance simultaneously using a GAM, in conmat

conmat is typically fitted to datasets *aggregated* across all pairs of ages or age groups. Since the contact rate is assumed to be the same for all these data they can be modelled as a single Poisson observation (with the number of participants given as an offset), which is significantly more computationally efficient than running the model across all pairs of participant ID and contact age and yields a model with the same likelihood. This is no longer possible in the combined model, since each participant has a distinct contact rate.

For the French Connection data, with 4 waves, we therefore have up to 400 (4 observed days, 100 contact ages) datapoints per participant, resulting in over 400,000 rows of data. Whereas the aggregated dataset used to fit only age effects above has just over 9,000 rows of data.

```{r}
# subset to required columns and add the number of participants per row (always 1 now)
fc_contact_data_participant <- fc_contact_data |>
  select(
    age_from,
    age_to,
    setting,
    contacts,
    part_id
  ) |>
  mutate(
    participants = 1
  )

nrow(fc_contact_data_participant)
```

Whilst we can estimate the model with the following code, using the `heterogeneous-augmentation` branch of conmat. By including a 'part_id' column in the data passed to conmat, an IID random effect on this variable will be included in the GAM. However this will take a very long time to fit.

```{r eval=FALSE}
fc_model_participant <- fit_single_contact_model(
  contact_data = fc_contact_data_participant,
  population = fc_population
)
```

We can reduce this computational complexity by reducing the resolution of the contact ages into bins, and using the midpoint age of the bin. This is an approximation to the likelihood (equivalent to the windowing method for approximating a Poisson process likelihood with a midpoint integration rule), and will reduce the age-resolution of the age-contact patterns estimated by contact, potentially over-smoothing the age-effect estimates. By reducing the contact age resolution into 5 year bins, we can reduce the number of datapoints for fitting to around 35,000.

```{r}
# set the maximum age in the data
max_age_data <- 99
resolution <- 5
age_agg_breaks <- c(seq(0, max_age_data, by = resolution), Inf)

# create a lookup to aggregate ages into 5 year bins
age_agg_lookup <- tibble(
  lower = age_agg_breaks[-length(age_agg_breaks)],
  upper = age_agg_breaks[-1]
  ) |>
  mutate(
    label = sprintf("[%s,%s)", lower, upper),
    upper = case_when(
      !is.finite(upper) ~ max_age_data + 1,
      .default = upper),
    midpoint = lower + 2
  ) |>
  rowwise() |>
  mutate(
    age = list(seq(lower, upper - 1, by = 1))
  ) |>
  unnest(age) |>
  select(
    age,
    label,
    midpoint
  )

fc_contact_data_participant_lores <- fc_contact_data_participant |>
  left_join(
    age_agg_lookup,
    by = c(age_to = "age")
  ) |>
  mutate(
    age_to_midpoint = midpoint
  ) |>
  group_by(
    part_id,
    age_from,
    setting,
    age_to_midpoint
  ) |>
  summarise(
    contacts = sum(contacts),
    .groups = "drop"
  ) |>
  mutate(
    participants = 1
  ) |>
  rename(
    age_to = age_to_midpoint
  )

nrow(fc_contact_data_participant_lores)
```

The following code runs the model on this lower resolution dataset, however it still takes an unreasonable amount of time to run.

```{r eval = FALSE}
# run conmat on this and return the age-structured matrix
fc_model_participant_lores <- fit_single_contact_model(
  contact_data = fc_contact_data_participant_lores,
  population = fc_population
)
```

We can compute and visualise the age-structured matrix estimated via this method

```{r eval=FALSE}
age_matrix_participant <- fc_model_participant_lores |>
  predict_contacts(
    population = fc_population,
    age_breaks = age_breaks
  ) |>
  predictions_to_matrix()

# output the discretised age matrix
age_matrix_participant_plot <- autoplot(age_matrix_participant)
ggsave("figures/fc_age_matrix_participant.png",
       plot = age_matrix_participant_plot,
       bg = "white",
       width = 6,
       height = 5,
       dpi = 300)
age_matrix_participant_plot
```

And finally, we can extract the standard deviation of the participant random effect.

```{r eval=FALSE}
variances <- mgcv::gam.vcomp(fc_model_participant_lores)
idx <- match("s(part_id)", names(variances))
sigma_fc_onestage <- sqrt(variances[idx])
sigma_fc_onestage
```

Now we have our estimate of the sigma parameter, we can construct an activity-structured contact matrix.

Then we can construct the age/activity contact matrix as the kronecker product of the age-structured matrix and the activity-structured matrices, and compute the population fractions for each combined population bin from the products of the fractions from the two models.

To construct an activity-structured contact matrix we need to select a number of bins to discretise the continuous activity distribution, and determine the parameters of the activity matrix - the degree of assortativity epsilon, and the range of assortativity along activity levels alpha. We use arbitrary values to begin with.

```{r}
n_activity_bins <- 4
alpha <- 0.5
epsilon <- 0.5
```

Next, we compute the discretisation bins of the continuous activity level distribution: the fraction of the population in each bin, and the average activity in the bin. For visulaisation here we use evenly-sized bins, so that the assortative patterns are clearly visible, rather than the contact rates being dominated by the different population fractions in each bin.

<!--# We do this using Gaussian quadrature, since it enables us to approximate the full activity distribution in relatively few bins: -->

```{r}
activity_bins <- quantile2_classes_mean(sigma = sigma_fc_twostage,
                                   n_classes = n_activity_bins)
activity_bins
```

Next we construct the fully-assortative activity-structured matrix, using Tom's iterative normalisation method and setting alpha as above:

```{r}
assortative_activity <- make_toms_contact_matrix(
  v = activity_bins$activity,
  p_pop = activity_bins$fraction,
  alpha = alpha)
```

And a proportionate mixing matrix, according to these activity levels and population fractions

```{r}
proportionate_activity <- make_proportionate_contact_matrix(
  v = activity_bins$activity,
  p_pop = activity_bins$fraction)
```

And now the final activity-structured matrix by blending the fully-assortative and fully-proportionate mixing matrices according to epsilon:

```{r}
activity_matrix <- make_blended_matrix(
  M_PM = proportionate_activity,
  M_assort = assortative_activity,
  eps = epsilon)
rownames(activity_matrix) <- activity_bins$class
colnames(activity_matrix) <- activity_bins$class
```

We can visuliase this

```{r}
activity_mat_plot <- autoplot_contact_matrix(activity_matrix, palette = 2) +
  ggtitle("activity-structured")
activity_mat_plot
```

Now we can combine the age and activity matrices into a combined age/activity structured matrix. For visualisation purposes, we make a lower-resolution version of the age-structured matrix, and use approximately evenly-sized age bins:

```{r}
ages <- fc_population$upper.age.limit
pops <- fc_population$population
cumpops <- cumsum(pops) / sum(pops)

# approximate the CDF of the age distribution and use this to estimate the integral to the max age
prob_max <- approx(x = ages,
                   y = cumpops,
                   xout = max_age,
                   rule = 2)$y

# approximate the quantile function of the age distribution, truncated at max_age by rescaling the cumulative densities, and use this to define even-sized breaks. 
cumpops_norm <- cumpops / prob_max

n_age_bins <- 6
breaks <- seq(0, 1, length.out = n_age_bins + 1)
age_breaks <- round(approx(x = cumpops_norm,
                           y = ages,
                           xout = breaks,
                           rule = 2)$y)

# force max age just in case of numerical errors
age_breaks[n_age_bins + 1] <- max_age
age_breaks[1] <- 0

# make contact matrix
age_contact_pred <- predict_contacts(fc_model_aggregated,
                                     fc_population,
                                     age_breaks = age_breaks)

age_matrix <- conmat::predictions_to_matrix(age_contact_pred)
age_mat_plot <- autoplot_contact_matrix(age_matrix, palette = 1) +
  ggtitle("age-structured")
age_mat_plot
```

And we combine this with our activity matrix

```{r}
age_activity_matrix <- tile_matrices(age_matrix, activity_matrix)
```

We can visualise these together

```{r}
age_activity_mat_plot <- autoplot_contact_matrix(age_activity_matrix, palette = 7) +
  ggtitle("age/activity-structured")
age_activity_mat_plot

```

Finally, combine the three illustrative matrices into a single three-panel figure (age-structured, activity-structured, age/activity-structured) for the manuscript.

```{r}
illustrative_matrices_plot <- age_mat_plot +
  activity_mat_plot +
  age_activity_mat_plot +
  plot_layout(nrow = 1)
ggsave("figures/illustrative_contact_matrices.png",
       plot = illustrative_matrices_plot,
       bg = "white",
       width = 15,
       height = 5,
       dpi = 300)
illustrative_matrices_plot
```

Next we recompute these matrices for a predetermined set of 5 year age bins, and a more optimal configuration of activity bins, and investigate the implications for analyses of disease transmission.

Compute the age matrix, and the population fraction in each of these age bins:

```{r}
age_breaks <- seq(0, max_age, by = 5)
age_contact_pred <- predict_contacts(fc_model_aggregated,
                                     fc_population,
                                     age_breaks = age_breaks)
age_matrix <- conmat::predictions_to_matrix(age_contact_pred)
pop_bin <- fc_population |>
  filter(lower.age.limit < max_age) |>
  pull(population)
age_fractions <- pop_bin / sum(pop_bin)
```

Activity matrix, with 10 bins and Gaussian quadrature for bin sizes and activity levels

```{r}
activity_matrix <- make_activity_matrix(n_activity_bins = 10,
                                        sigma = sigma_fc_twostage,
                                        alpha = 0.5,
                                        epsilon = 0.5)
activity_fractions <- attr(activity_matrix, "activity_bins")$fraction
```

Combined matrix, and activity_fractions

```{r}
age_activity_matrix <- tile_matrices(age_matrix, activity_matrix)
age_activity_fractions <- tile_fractions(age_fractions, activity_fractions)
```

Compute the R0 of next-generation matrices for the age-only and the age-activity contact matrices, given a per-contact transmission probability (beta) of 0.1.

```{r}
beta <- 0.1
R0_age <- get_eigenval(age_matrix * beta)
R0_age_activity <- get_eigenval(age_activity_matrix * beta)

R0_age
R0_age_activity
```

Now see how they differ in final size calculations. Compute the final size for the two models above, and for an age-only contact matrix with R0 scaled up to the same R0 as the age-activity NGM.

```{r}
# mock up susceptibility and p_susceptibility matrices (all set to 1, since we are starting at a naive population)
n_age <- length(age_fractions)
n_age_activity <- length(age_activity_fractions)
dummy_age <- as.matrix(rep(1, n_age))
dummy_age_activity <- as.matrix(rep(1, n_age_activity))

# scale the contact matrices, so that when multiplied by age fractions they have an eigenvalue of exactly 1 (an undocumented requirement of final_size)
age_matrix_scaled <- age_matrix / get_eigenval(age_matrix * age_fractions)
age_activity_matrix_scaled <- age_activity_matrix / get_eigenval(age_activity_matrix * age_activity_fractions)

# age contact matrix with age R0
size_age <- final_size(r0 = R0_age,
                       contact_matrix = age_matrix_scaled,
                       demography_vector = age_fractions,
                       susceptibility = dummy_age,
                       p_susceptibility = dummy_age)

# age contact matrix with age/activity R0
size_age_scaled <- final_size(R0_age_activity,
                       contact_matrix = age_matrix_scaled,
                       demography_vector = age_fractions,
                       susceptibility = dummy_age,
                       p_susceptibility = dummy_age)

# age/activity contact matrix with age/activity R0
size_age_activity <- final_size(R0_age_activity,
                       contact_matrix = age_activity_matrix_scaled,
                       demography_vector = age_activity_fractions,
                       susceptibility = dummy_age_activity,
                       p_susceptibility = dummy_age_activity)
```

Collapse the age-activity final sizes down into age groups

```{r}
size_age_activity_by_age <- size_age_activity |>
  mutate(
    demo_grp = str_split_i(demo_grp, "-", 1),
    ) |>
  group_by(
    demo_grp
  ) |>
  summarise(
    p_infected = sum(p_infected * activity_fractions))
```

## Bootstrap uncertainty

To quantify sampling uncertainty in these estimates we use a participant-level (cluster) bootstrap: we resample the survey participants with replacement (giving each drawn participant a fresh identifier, so a participant drawn more than once contributes as distinct individuals) and re-run the entire two-stage pipeline - age-structured contact model, residual heterogeneity `sigma`, activity matrix, R0s and final sizes - on each resample. The contact ages are re-imputed for each replicate, so imputation uncertainty is folded in as well.

This is computationally expensive, so the bootstrap is run separately by `R/bootstrap_analysis_3.R` (which parallelises across cores and caches its output). Here we load the cached replicates and summarise them. We report **bias-corrected ("BC")** bootstrap intervals: percentile intervals adjusted for any median bias between the point estimate and the bootstrap distribution. The acceleration term of a full BCa interval is deferred - it needs either a separate jackknife or a jackknife-after-bootstrap from the stored resample membership, and is only worth adding if the bootstrap distributions turn out to be materially skewed (inspect `z0` in the summary table below).

```{r}
boot <- readRDS("data/boot_analysis_3.rds")

# keep only the replicates that fitted successfully
boot_ok <- boot$replicates |>
  keep(~ is.null(.x$result$error))

n_boot <- length(boot_ok)

# tidy by-age final sizes across replicates (raw scenario keys)
boot_by_age <- boot_ok |>
  map_dfr(~ mutate(.x$result$by_age, rep = .x$rep))

# tidy scalar quantities across replicates
boot_scalars <- boot_ok |>
  map_dfr(~ tibble(
    rep = .x$rep,
    sigma = .x$result$sigma,
    R0_age = .x$result$R0_age,
    R0_age_activity = .x$result$R0_age_activity,
    overall_age = .x$result$overall[["age"]],
    overall_age_scaled = .x$result$overall[["age_scaled"]],
    overall_age_activity = .x$result$overall[["age_activity"]]
  ))

sprintf("%i of %i bootstrap replicates used (%i failed)",
        n_boot, boot$config$B, boot$config$B - n_boot)
```

Plot these age distributions

```{r}
# scenario labels, including the actual R0 each final size was computed under
# (ordered as they should appear in the legend)
scenario_labels <- c(
  age          = sprintf("age matrix (R0 = %.2f)", R0_age),
  age_activity = sprintf("age/activity matrix (R0 = %.2f)", R0_age_activity),
  age_scaled   = sprintf("age matrix, rescaled (R0 = %.2f)", R0_age_activity)
)

# combine the three scenarios into one tibble, and parse the lower/upper edges
# of each age bin from the group label (e.g. "[0,5)") so the bins can be placed
# numerically on the x axis
final_size_scenarios <- bind_rows(
  age          = size_age,
  age_scaled   = size_age_scaled,
  age_activity = size_age_activity_by_age,
  .id = "scenario_key"
) |>
  mutate(
    scenario = factor(
      scenario_labels[scenario_key],
      levels = scenario_labels
    ),
    age_lower = as.numeric(str_match(demo_grp, "([0-9]+)[^0-9]+([0-9]+)")[, 2]),
    age_upper = as.numeric(str_match(demo_grp, "([0-9]+)[^0-9]+([0-9]+)")[, 3])
  ) |>
  arrange(scenario, age_lower)

# bootstrap BC interval for each scenario x age bin, to draw as a band. The band
# is piecewise-constant across each age bin (a geom_rect per bin), matching the
# step lines exactly (a geom_ribbon would slope between bin midpoints instead).
final_size_ci <- final_size_scenarios |>
  select(scenario_key, scenario, age_lower, age_upper, p_infected) |>
  pmap_dfr(function(scenario_key, scenario, age_lower, age_upper, p_infected) {
    reps <- boot_by_age |>
      filter(scenario == scenario_key, age_lower == !!age_lower) |>
      pull(p_infected)
    bc_ci(reps, p_infected) |>
      mutate(
        scenario = scenario,
        age_lower = age_lower,
        age_upper = age_upper
      )
  })

# add a terminal point at the upper edge of the final bin for each scenario, so
# the step spans the full width of every age bin (piecewise constant)
final_size_steps <- final_size_scenarios |>
  group_by(scenario) |>
  group_modify(~ bind_rows(
    .x,
    mutate(slice_tail(.x, n = 1), age_lower = age_upper)
  )) |>
  ungroup()

# points at the bin midpoints, to carry the per-scenario symbols
final_size_points <- final_size_scenarios |>
  mutate(age_mid = (age_lower + age_upper) / 2)

final_size_plot <- ggplot() +
  # 95% bootstrap BC interval, piecewise-constant across each age bin
  geom_rect(
    data = final_size_ci,
    aes(xmin = age_lower, xmax = age_upper,
        ymin = bc_lower, ymax = bc_upper, fill = scenario),
    alpha = 0.2
  ) +
  geom_step(
    data = final_size_steps,
    aes(x = age_lower, y = p_infected, colour = scenario),
    direction = "hv",
    linewidth = 0.8
  ) +
  scale_colour_brewer(palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2") +
  guides(fill = "none") +
  scale_y_continuous(labels = scales::label_percent()) +
  coord_cartesian(ylim = c(0, 1)) +
  labs(
    x = "age",
    y = "final attack rate",
    colour = "scenario"
  ) +
  theme_minimal()

ggsave("figures/final_size_by_age.png",
       plot = final_size_plot,
       bg = "white",
       width = 7,
       height = 5,
       dpi = 300)

final_size_plot
```

Compute the overall (population-weighted) final size from these bin-specific final sizes

```{r}
overall_size_age <- sum(size_age$p_infected * age_fractions)
overall_size_age_scaled <- sum(size_age_scaled$p_infected * age_fractions)
overall_size_age_activity <- sum(size_age_activity$p_infected * age_activity_fractions)
```

Finally, assemble the headline quantities with their point estimates and 95% bootstrap intervals. We show the bias-corrected ("BC") interval alongside the plain percentile interval, and the bias-correction constant `z0` (large `|z0|` would indicate the bootstrap distribution is noticeably off-centre from the point estimate, and hence that adding the BCa acceleration term might be worthwhile).

```{r}
summary_table <- bind_rows(
  bc_ci(boot_scalars$R0_age, R0_age) |>
    mutate(quantity = "R0 (age matrix)"),
  bc_ci(boot_scalars$R0_age_activity, R0_age_activity) |>
    mutate(quantity = "R0 (age/activity matrix)"),
  bc_ci(boot_scalars$sigma, sigma_fc_twostage) |>
    mutate(quantity = "residual heterogeneity (sigma)"),
  bc_ci(boot_scalars$overall_age, overall_size_age) |>
    mutate(quantity = "overall final size (age matrix)"),
  bc_ci(boot_scalars$overall_age_scaled, overall_size_age_scaled) |>
    mutate(quantity = "overall final size (age matrix, rescaled)"),
  bc_ci(boot_scalars$overall_age_activity, overall_size_age_activity) |>
    mutate(quantity = "overall final size (age/activity matrix)")
) |>
  transmute(
    quantity,
    estimate,
    `BC 95% CI` = sprintf("(%.3f, %.3f)", bc_lower, bc_upper),
    `percentile 95% CI` = sprintf("(%.3f, %.3f)", pct_lower, pct_upper),
    z0 = round(z0, 2)
  )

knitr::kable(summary_table, digits = 3)
```
