---
title: "Analysis - Continuous contact model"
format: html
editor: visual
---

# Continuous kernel-based contact model

This document is an R translation of `Matlab/contact_models.m`. It visualises contact matrices and dominant eigenvalues under two continuous-quantile-space assortativity models:

-   **Mike's method** (`make_assortative_contact_matrix`): iterative column-by-column construction.
-   **Tom's method** (`make_toms_contact_matrix`): iterative kernel normalisation.

Both methods parametrise assortativity via a Gaussian kernel with width parameter `b`. Larger `b` concentrates mixing among individuals with similar activity levels; `b = 0` recovers proportionate mixing. An epsilon parameter blends the assortative matrix with the proportionate mixing baseline.

```{r}
#| message: false
#| warning: false
source("R/packages.R")
source("R/functions.R")
```

## Setup

Set up the quantile grid and activity level distribution (mirrors `contact_models.m` lines 1–55).

```{r}
# Model parameters
sigma        <- 0.4             # fixed sigma for (eps, b) sweep
eps_arr      <- seq(0, 1, by = 0.05)
b_arr        <- seq(0, 40, by = 2)
sigma_arr    <- seq(0, 1, by = 0.1)

# Indices (1-based) of (eps, b) combinations to show in matrix grid plots
eps_to_plot <- c(1, 11, 21)
b_to_plot   <- c(1, 11, 21)

# Quantile grid
n_bins <- 100
dx     <- 1 / n_bins
x      <- seq(dx / 2, 1 - dx / 2, by = dx)

# Activity level distribution: inverse CDF of lognormal(0, sigma) at each quantile
v  <- qlnorm(x, meanlog = 0, sdlog = sigma)
Ev <- dx * sum(v)

# Meshgrid matrices for matrix construction:
#   X[i, j] = x[j]  (individual's activity quantile, varies along columns)
#   Y[i, j] = x[i]  (contact's activity quantile, varies along rows)
X <- matrix(x, nrow = n_bins, ncol = n_bins, byrow = TRUE)
Y <- matrix(x, nrow = n_bins, ncol = n_bins, byrow = FALSE)

# Proportionate mixing matrix (equivalent to MATLAB: dx * v' .* v / Ev)
M_PM <- dx * outer(v, v) / Ev
```

## Sweep over (epsilon, b)

Compute the dominant eigenvalue for a grid of (epsilon, b) values using both methods. Run in parallel.

```{r}
plan(multisession, workers = 8)

sweep_eps_b <- expand_grid(
  eps = eps_arr,
  b   = b_arr
) |>
  mutate(
    eigenval_mike = furrr::future_map2(
      eps, b,
      \(e, bval) {
        M_assort <- make_assortative_contact_matrix(X, Y, v, bval)
        get_eigenval(make_blended_matrix(M_PM, M_assort, e))
      },
      .options = furrr_options(seed = TRUE)
    ),
    eigenval_tom = furrr::future_map2(
      eps, b,
      \(e, bval) {
        T_assort <- make_toms_contact_matrix(x, v, bval, dx)
        get_eigenval(make_blended_matrix(M_PM, T_assort, e))
      },
      .options = furrr_options(seed = TRUE)
    )
  ) |>
  unnest(c(eigenval_mike, eigenval_tom))
```

## Sweep over (sigma, b) with epsilon = 1

```{r}
sweep_sigma_b <- expand_grid(
  sigma_val = sigma_arr,
  b         = b_arr
) |>
  mutate(
    eigenval_mike = furrr::future_map2(
      sigma_val, b,
      \(s, bval) {
        v_s <- qlnorm(x, meanlog = 0, sdlog = s)
        # qlnorm with sdlog = 0 returns 1 for all x in (0, 1); no special case needed
        Ev_s  <- dx * sum(v_s)
        M_PM_s <- dx * outer(v_s, v_s) / Ev_s
        M_assort <- make_assortative_contact_matrix(X, Y, v_s, bval)
        get_eigenval(M_assort)   # eps = 1: pure assortative
      },
      .options = furrr_options(seed = TRUE)
    ),
    eigenval_tom = furrr::future_map2(
      sigma_val, b,
      \(s, bval) {
        v_s <- qlnorm(x, meanlog = 0, sdlog = s)
        T_assort <- make_toms_contact_matrix(x, v_s, bval, dx)
        get_eigenval(T_assort)
      },
      .options = furrr_options(seed = TRUE)
    )
  ) |>
  unnest(c(eigenval_mike, eigenval_tom))
```

## Figure 1 & 2: Contact matrix grids

Visualise the contact matrix for selected (eps, b) combinations (3 × 3 grid). Rebuild matrices for the 9 selected parameter pairs.

```{r}
plot_params <- expand_grid(
  eps = eps_arr[eps_to_plot],
  b   = b_arr[b_to_plot]
)

matrix_long_mike <- pmap(
  plot_params,
  \(eps, b) {
    M_assort <- make_assortative_contact_matrix(X, Y, v, b)
    M_blend  <- make_blended_matrix(M_PM, M_assort, eps)
    expand_grid(y_val = x, x_val = x) |>
      mutate(
        M_val = as.vector(M_blend),
        eps   = eps,
        b     = b
      )
  }
) |>
  list_rbind()

matrix_long_mike |>
  ggplot(aes(x = x_val, y = y_val, fill = M_val)) +
  geom_raster() +
  scale_fill_viridis_c(trans = "log") +
  facet_grid(
    rows = vars(paste0("eps = ", eps)),
    cols = vars(paste0("b = ", b))
  ) +
  labs(
    x    = "activity level quantile of individual (x)",
    y    = "activity level quantile of contact (y)",
    fill = "M(x, y)",
    title = "Mike's method: contact matrix for selected (eps, b)"
  ) +
  theme_minimal()
```

```{r}
matrix_long_tom <- pmap(
  plot_params,
  \(eps, b) {
    T_assort <- make_toms_contact_matrix(x, v, b, dx)
    T_blend  <- make_blended_matrix(M_PM, T_assort, eps)
    expand_grid(y_val = x, x_val = x) |>
      mutate(
        M_val = as.vector(T_blend),
        eps   = eps,
        b     = b
      )
  }
) |>
  list_rbind()

matrix_long_tom |>
  ggplot(aes(x = x_val, y = y_val, fill = M_val)) +
  geom_raster() +
  scale_fill_viridis_c(trans = "log") +
  facet_grid(
    rows = vars(paste0("eps = ", eps)),
    cols = vars(paste0("b = ", b))
  ) +
  labs(
    x    = "activity level quantile of individual (x)",
    y    = "activity level quantile of contact (y)",
    fill = "M(x, y)",
    title = "Tom's method: contact matrix for selected (eps, b)"
  ) +
  theme_minimal()
```

## Figure 3: Aggregate contact profiles

For each selected (eps, b) combination, compare the column sums of the blended matrix (= total contact rate as a function of activity quantile) against the proportionate mixing baseline and the underlying log-normal activity distribution.

```{r}
agg_long <- pmap(
  plot_params,
  \(eps, b) {
    M_assort <- make_assortative_contact_matrix(X, Y, v, b)
    M_blend  <- make_blended_matrix(M_PM, M_assort, eps)
    T_assort <- make_toms_contact_matrix(x, v, b, dx)
    T_blend  <- make_blended_matrix(M_PM, T_assort, eps)
    tibble(
      x       = x,
      PM      = colSums(M_PM),
      Mike    = colSums(M_blend),
      Tom     = colSums(T_blend),
      lognorm = v,
      eps     = eps,
      b       = b
    )
  }
) |>
  list_rbind() |>
  pivot_longer(
    cols      = c(PM, Mike, Tom, lognorm),
    names_to  = "series",
    values_to = "value"
  )

agg_long |>
  ggplot(aes(x = x, y = value, colour = series, linetype = series)) +
  geom_line() +
  facet_grid(
    rows = vars(paste0("eps = ", eps)),
    cols = vars(paste0("b = ", b))
  ) +
  labs(
    x      = "activity level quantile",
    y      = "total contact rate",
    colour = NULL,
    linetype = NULL,
    title  = "Column sums of blended matrix vs. activity distribution"
  ) +
  theme_minimal()
```

## Figure 4: Eigenvalue landscapes

### Contour plot: eigenvalue over (b, epsilon)

```{r}
sweep_eps_b |>
  pivot_longer(
    cols      = c(eigenval_mike, eigenval_tom),
    names_to  = "method",
    values_to = "eigenval"
  ) |>
  mutate(
    method = recode(
      method,
      eigenval_mike = "Mike's method",
      eigenval_tom  = "Tom's method"
    )
  ) |>
  ggplot(aes(x = b, y = eps, z = eigenval)) +
  geom_contour_filled(breaks = seq(1.1, 1.8, by = 0.05)) +
  facet_wrap(~ method) +
  labs(
    x    = "b (kernel width)",
    y    = expression(epsilon ~ "(assortativity blend)"),
    fill = expression(lambda),
    title = sprintf("Dominant eigenvalue over (b, epsilon) [sigma = %.1f]", sigma)
  ) +
  theme_minimal()
```

### Eigenvalue vs sigma for selected b values

```{r}
b_pick <- b_arr[seq(1, length(b_arr), by = 5)]

sweep_sigma_b |>
  filter(b %in% b_pick) |>
  pivot_longer(
    cols      = c(eigenval_mike, eigenval_tom),
    names_to  = "method",
    values_to = "eigenval"
  ) |>
  mutate(
    method = recode(
      method,
      eigenval_mike = "Mike's method",
      eigenval_tom  = "Tom's method"
    )
  ) |>
  ggplot(
    aes(
      x      = sigma_val,
      y      = eigenval,
      colour = factor(b),
      group  = factor(b)
    )
  ) +
  geom_line() +
  facet_wrap(~ method) +
  labs(
    x      = expression(sigma ~ "(activity level SD)"),
    y      = "dominant eigenvalue",
    colour = "b",
    title  = "Dominant eigenvalue vs. sigma for selected b values (eps = 1)"
  ) +
  theme_minimal()
```
