---
title: "Analysis 1 - variance_partitioning"
format: html
editor: visual
---

## Variance partitioning analysis on contact rates

How much variation in contact rates are we missing?

#### Analyse partitioning of variation in contact rates (within/between individuals and strata)

Load packages and functions

```{r}
#| message: false
#| warning: false
source("R/packages.R")
source("R/functions.R")

# set the RNG seed - used by the permutation check below
set.seed(2026-07-07)
```

Load the processed French Connection and Hong Kong data. We add an observation identifier `obs` (one level per participant/wave/day) so that we can fit an observation-level random effect (see below).

```{r}
fc_contact_counts <- process_fc_data_varpart() |>
  mutate(obs = factor(row_number()))
hk_contact_counts <- process_hk_data_varpart() |>
  mutate(obs = factor(row_number()))
```

We model the daily contact counts as Poisson with a log link (consistent with the other analyses in this repo). The random effects capture the within-individual variation (between observed days), the between-individual variation explained by age, and the residual between-individual variation.

A Poisson-log GLMM has no residual variance parameter: the day-to-day variation within an individual is otherwise assumed to be exactly Poisson sampling noise. Contact counts are overdispersed, so we add an **observation-level random effect** `(1|obs)` to capture genuine within-individual, extra-Poisson variation. Without it, that overdispersion would inflate the between-individual (`part_id`) variance. We fit with `nAGQ = 0` for numerical robustness, consistent with the two-stage model in analysis 3.

First we fit for the France data.

```{r}
#| message: false
#| warning: false

fc_model_none <- glmer(
  contacts ~ (1|part_id) + (1|obs),
  family = poisson,
  data = fc_contact_counts,
  nAGQ = 0
)

fc_model_age <- glmer(
  contacts ~ (1|part_age) +
    (1|part_id) + (1|obs),
  family = poisson,
  data = fc_contact_counts,
  nAGQ = 0
)

# the same but with gender and occupation (confounded)
fc_model_all <- glmer(
  contacts ~ (1|part_age) +
    (1|part_gender) +
    (1|part_occupation) +
    (1|part_id) + (1|obs),
  family = poisson,
  data = fc_contact_counts,
  nAGQ = 0
)
```

Now for Hong Kong

```{r}
#| message: false
#| warning: false
hk_model_none <- glmer(
  contacts ~ (1|part_id) + (1|obs),
  family = poisson,
  data = hk_contact_counts,
  nAGQ = 0
)

hk_model_age <- glmer(
  contacts ~ (1|part_age) +
    (1|part_id) + (1|obs),
  family = poisson,
  data = hk_contact_counts,
  nAGQ = 0
)

# the same but with gender (confounded)
hk_model_all <- glmer(
  contacts ~ (1|part_age) +
    (1|part_gender) +
    (1|part_id) + (1|obs),
  family = poisson,
  data = hk_contact_counts,
  nAGQ = 0
)
```

Do variance partitioning on each, combining across studies

```{r}
partitioning_none <- bind_rows(
  France = partition_variance_glmer(fc_model_none),
  `Hong Kong` = partition_variance_glmer(hk_model_none),
  .id = "Study"
)

partitioning_age <- bind_rows(
  France = partition_variance_glmer(fc_model_age),
  `Hong Kong` = partition_variance_glmer(hk_model_age),
  .id = "Study"
)

partitioning_all <- bind_rows(
  France = partition_variance_glmer(fc_model_all),
  `Hong Kong` = partition_variance_glmer(hk_model_all),
  .id = "Study"
)
```

Arrange them in a multipanel figure, for supplemental materials

```{r}
# panel C ("additional predictors") is the only panel containing all five
# components, so let it carry the legend (its keys all have data and render
# correctly); suppress the legend on A and B so a single, shared legend is
# collected for the whole figure
panel_age <- make_barplot(partitioning_age) +
  ggtitle("Only age variable") +
  theme(legend.position = "none")
panel_none <- make_barplot(partitioning_none) +
  ggtitle("No demographic variables") +
  theme(legend.position = "none") +
  # match the x-axis padding panel C gets from its leader-line labels, so the
  # bars in panels B and C have the same width
  scale_x_discrete(expand = expansion(add = 0.9))
# raise the threshold so the (small) between-ages slice is labelled with a
# leader line too, rather than crammed inside the bar next to between-genders
panel_all <- make_barplot(partitioning_all, label_threshold = 0.07) +
  ggtitle("All demographic variables")

# panel A (age only) large on the left; B (no covariates) and C (all covariates)
# stacked on the right
design <- "
AB
AC
"

varpart_all <- panel_age + panel_none + panel_all +
  plot_layout(design = design, guides = "collect") +
  plot_annotation(
    # title = "Partitioning variance in daily log contact rates",
    tag_levels = "A"
  )

ggsave("figures/varpart_all.png",
       plot = varpart_all,
       bg = "white",
       scale = 1,
       width = 9,
       height = 6)
```

Tabulate the same percentages (formatted as in the figure) for each model and study.

```{r}
component_levels <- c(
  "within individuals",
  "between individuals (unexplained)",
  "between ages",
  "between genders",
  "between occupations"
)
model_levels <- c(
  "No demographic variables",
  "Only age variable",
  "All demographic variables"
)

partitioning_all_models <- bind_rows(
  `No demographic variables` = partitioning_none,
  `Only age variable` = partitioning_age,
  `All demographic variables` = partitioning_all,
  .id = "Model"
)

# one row per study/model, one column per component, cells as formatted
# percentages (a dash where a component is not in the model)
partition_table <- function(data, levels) {
  data |>
    mutate(
      Model = factor(Model, levels = model_levels),
      partition = factor(partition, levels = levels),
      percent = format_percent_label_1dp(proportion)
    ) |>
    select(Study, Model, partition, percent) |>
    pivot_wider(
      names_from = partition,
      values_from = percent,
      names_expand = TRUE
    ) |>
    arrange(Study, Model) |>
    mutate(across(where(is.character), ~ replace_na(.x, "--")))
}

# write a human-readable text (markdown) version to tables/, and display it
write_table_txt <- function(tab, caption, path) {
  writeLines(c(caption, "", knitr::kable(tab, format = "pipe")), path)
}

caption_full <- "Variance in daily log contact rates (%), partitioned by component, for each model and study."
partition_full <- partition_table(partitioning_all_models, component_levels)
write_table_txt(partition_full, caption_full, "tables/variance_partition.txt")
knitr::kable(partition_full, caption = caption_full)
```

The same partition after excluding the within-individual component, with the remaining (between-individual) components renormalised to sum to 100%.

```{r}
caption_between <- "Between-individual variance (%), partitioned among the age/gender/occupation and residual components (i.e. excluding within-individual variance and renormalising to 100%), for each model and study."
partition_between <- partitioning_all_models |>
  filter(partition != "within individuals") |>
  group_by(Study, Model) |>
  mutate(proportion = proportion / sum(proportion)) |>
  ungroup() |>
  partition_table(component_levels[-1])
write_table_txt(partition_between, caption_between, "tables/variance_partition_excluding_within.txt")
knitr::kable(partition_between, caption = caption_between)
```

Sanity check this analysis by permuting the participant IDs - we would expect the proportion of variance within individuals (not explained by age/individual) to be high if we shuffled the data around between individuals/repeats. So confirm this.

```{r}
#| message: false
#| warning: false
# sanity check contact variance analysis by permuting the participant IDs
permute_sim <- function() {
  fc_contact_counts %>%
    mutate(
      contacts = sample(contacts)
    ) %>%
    glmer(
      contacts ~ (1|part_age) + (1|part_id) + (1|obs),
      family = poisson,
      data = .,
      nAGQ = 0
    ) %>%
    partition_variance_glmer() %>%
    select(-var) %>%
    pivot_wider(
      names_from = "partition",
      values_from = "proportion"
    )
}

set.seed(2026-07-07)
sims_list <- replicate(50,
                       permute_sim(),
                  simplify = FALSE)
sims_df <- do.call(bind_rows, sims_list)

partitions <- partitioning_age$partition

for (i in seq_along(partitions)) {
  this_partition <- partitions[i]
  this_study <- partitioning_age$Study[i]
  perturbed_props <- sims_df[[this_partition]]
  estimated_prop <- partitioning_age |>
    filter(partition == this_partition) |>
    pull(proportion)
  
  xlim <- range(c(perturbed_props, estimated_prop))
  
  hist(perturbed_props,
       xlim = xlim,
       breaks = 100,
       xlab = "Proportion",
       main = paste(this_study,
                    this_partition,
                    sep = "\n"))
  abline(v = estimated_prop)
}
```

Extract the between-individual standard deviation (`part_id`) from the age model. Because the model uses a log link, this is a standard deviation of individual log contact rates. This is comparable to the `sigma` of a lognormal individual activity distribution, as used in the full model. Though note that here we not account for heterogeneity in the contact ages when computing this parameter, so this is not intended as an estimate of `sigma` in that model.

```{r}
extract_part_id_sd <- function(model) {
  model |>
    summary() |>
    extracting("varcor") |>
    as_tibble() |>
    filter(grp == "part_id") |>
    pull(sdcor)
}

fc_sigma <- extract_part_id_sd(fc_model_age)
hk_sigma <- extract_part_id_sd(hk_model_age)

fc_sigma
hk_sigma
```
