---
title: "Analysis S1 - Binning methods"
format: html
editor: visual
---

# Comparing binning schemes for summarising the activity level distribution

The computational efficiency of the proposed method in a real-world analysis relies on the ability to represent the population distribution of activity levels in a comparatively low number of bins. Specifically, we aim to rapidly converge on the true eigenvalue for a population, as the number of bins increases. We find that defining bin sizes and activity levels according to a Gaussian quadrature scheme achieves a good level of convergence, and outperforms using evenly-sized bins and setting the bin activity level to the bin mean.

Here we illustrate the differences between these two alternative binning methods, both visually and in terms of representing the mean and variance of the distribution, and then demonstrate that the Gaussian quadrature scheme converges on the true eigenvalue considerably more quickly than the evenly-sized bins approach.

The three binning methods we consider are defined as follows

#### Quantiles with midpoint activity

Define bins of equal population size by defining evenly-spaced quantiles of the distribution over activity levels, and set the activity level for each bin to the value of the distribution at the bin's midpoint quantile (the inverse cumulative distribution function evaluated at that midpoint).

#### Quantiles with mean activity

Define bins of equal population size in the same way (up to some maximum quantile less than 1), but set the activity level for each bin to its mean, computed by numerical integration via one dimensional adaptive quadrature over the activity range of the bin, rather than to the bin midpoint value.

#### Gaussian quadrature scheme

Define bins of unequal size according to the nodes of a Guassian quadrature scheme, with activity levels given by the node locations, and population fractions given by the normalised weights.

```{r echo = FALSE}
#| message: false
#| warning: false
source("R/packages.R")
source("R/functions.R")
```

#### Illustrating the binning schemes

First, we illustrate how the binning schemes place their bins relative to the target distribution. For each scheme we use 7 bins and set $\sigma = 0.5$. The top-left panel shows the target lognormal density of activity levels, corresponding to our model of activity. The remaining panels show the bin activity levels (points) and population fractions (vertical stems) for the quantile-midpoint, quantile-mean, and Gaussian quadrature schemes respectively.

```{r}
#| message: false
#| warning: false
sigma <- 0.5
n_int <- 7

# quantile mean, quantile midpoint, and Gaussian quadrature classes
quantile_classes <- quantile2_classes_mean(sigma = sigma,
                                            n_classes = n_int)
midpoint_classes <- quantile_classes_midpoint(sigma = sigma,
                                              n_classes = n_int)
gq_classes <- gaussquad_classes(sigma = sigma,
                                n_classes = n_int)

# common x-axis range
xmax <- max(
  qlnorm(0.99, 0, sigma),
  max(quantile_classes$activity),
  max(midpoint_classes$activity),
  max(gq_classes$activity)
)

# top panel: target distribution
p_target <- ggplot() +
  geom_function(
    fun = dlnorm,
    args = list(meanlog = 0, sdlog = sigma)
  ) +
  xlim(0, xmax) +
  labs(
    title = "Target distribution",
    x = expression("activity level (" * italic(v) * ")"),
    y = expression("p(" * italic(v) * ")")
  ) +
  theme_minimal()

# middle panel: quantile mean classes
p_quantile <- ggplot(quantile_classes,
                     aes(x = activity, y = fraction)) +
  geom_segment(
    aes(xend = activity, yend = 0)
  ) +
  geom_point(size = 2) +
  xlim(0, xmax) +
  ylim(0, max(quantile_classes$fraction) * 1.1) +
  labs(
    title = "Quantile mean",
    x = expression("activity level (" * italic(v) * ")"),
    y = "population fraction"
  ) +
  theme_minimal()

# middle panel: quantile midpoint classes
p_midpoint <- ggplot(midpoint_classes,
                    aes(x = activity, y = fraction)) +
  geom_segment(
    aes(xend = activity, yend = 0)
  ) +
  geom_point(size = 2) +
  xlim(0, xmax) +
  ylim(0, max(midpoint_classes$fraction) * 1.1) +
  labs(
    title = "Quantile midpoint",
    x = expression("activity level (" * italic(v) * ")"),
    y = "population fraction"
  ) +
  theme_minimal()

# lower panel: Gaussian quadrature classes
p_gq <- ggplot(gq_classes,
              aes(x = activity, y = fraction)) +
  geom_segment(
    aes(xend = activity, yend = 0)
  ) +
  geom_point(size = 2) +
  xlim(0, xmax) +
  ylim(0, max(gq_classes$fraction) * 1.1) +
  labs(
    title = "Gaussian quadrature",
    x = expression("activity level (" * italic(v) * ")"),
    y = "population fraction"
  ) +
  theme_minimal()

binning_illustration <- (p_target | p_midpoint) / (p_quantile | p_gq) +
  plot_annotation(tag_levels = "A")

ggsave("figures/binning_illustration.png",
       plot = binning_illustration,
       bg = "white",
       scale = 1,
       width = 9,
       height = 6)
```

![Placement of bins by the three schemes relative to the target distribution, using 7 bins and $\sigma = 0.5$. (A) the target lognormal density of activity levels. (B--D) the bin activity levels (points) and population fractions (vertical stems) for the quantile midpoint (B), quantile mean (C), and Gaussian quadrature (D) schemes respectively.](../figures/binning_illustration.png){#fig-binning-illustration}

The Gaussian quadrature scheme is able to place bins much further into the two tails of the distribution than either evenly-sized-bin scheme (quantile midpoint and quantile mean). Whilst few individuals would have activity levels in these tails, those that do (particularly at the upper tail) contribute significantly to the population eigenvalue. The two evenly-sized-bin schemes place their bins identically; they differ only in the representative activity level assigned to each bin (the bin midpoint value versus the bin mean), with the midpoint value falling below the mean within each bin for this right-skewed distribution.

#### Approximating the distribution

Next we assess the ability of each binning scheme to approximate the first two moments of the true distribution, by computing numerical approximations to these mean and variance of the distribution over activity levels using each of the binning schemes as a quadrature scheme. We vary the number of bins or each scheme and compare the approximations to the mean and variance against the truth, for three values of sigma.

```{r}
estimates <- expand_grid(
  n_classes = 2:20,
  sigma = c(0.5, 1, 2),
  method = c("truth",
             "quantile_midpoint",
             "quantile2_mean",
             "gaussquad")
) |>
  rowwise() |>
  mutate(
    mean = case_when(
      method == "truth" ~ exp(0 + sigma ^ 2 / 2),
      .default = bin_estimate_mean(sigma,
                                   n_classes,
                                   method = method)
    ),
    sd = case_when(
      method == "truth" ~ sqrt((exp(sigma ^ 2) - 1) * exp(2 * 0 + sigma ^ 2)),
      .default = bin_estimate_sd(sigma,
                                 n_classes,
                                 method = method)
    )
  ) |>
  ungroup() |>
  # relabel methods for display (quantile2_mean is the deterministic
  # numerical-integration version of the quantile mean method)
  mutate(
    method = recode(
      method,
      quantile_midpoint = "quantile midpoint",
      quantile2_mean = "quantile mean",
      gaussquad = "Gaussian quadrature"
    )
  )

```

```{r}
# colour used for the Gaussian quadrature line. Methods are plotted in the order
# quantile midpoint, quantile mean, Gaussian quadrature (so Gaussian quadrature
# is drawn last, on top of the others), making it the third of the default hues
gq_colour <- scales::hue_pal()(3)[3]

# method display order: Gaussian quadrature last so its line is drawn on top
method_levels <- c("quantile midpoint", "quantile mean", "Gaussian quadrature")

moment_long <- estimates |>
  pivot_longer(
    cols = c("mean", "sd"),
    values_to = "estimate",
    names_to = "moment"
  )

# smallest number of bins for which the Gaussian quadrature estimate is within
# 1% of the true value, for each moment and value of sigma
moment_threshold <- moment_long |>
  filter(method == "Gaussian quadrature") |>
  left_join(
    moment_long |>
      filter(method == "truth") |>
      select(n_classes, sigma, moment, true = estimate),
    by = c("n_classes", "sigma", "moment")
  ) |>
  filter(abs(estimate - true) / true < 0.01) |>
  group_by(moment, sigma) |>
  summarise(
    bins_1pct = min(n_classes),
    .groups = "drop"
  ) |>
  mutate(
    sigma = paste0("sigma == ", sigma)
  )

# the true value of each moment, for each sigma
moment_truth <- moment_long |>
  filter(method == "truth") |>
  distinct(moment, sigma, true = estimate) |>
  mutate(
    sigma = paste0("sigma == ", sigma)
  )

# one panel letter per facet, in the facet_wrap reading order (mean row A-C,
# then sd row D-F)
panel_tags_moment <- moment_long |>
  distinct(moment, sigma) |>
  arrange(moment, sigma) |>
  mutate(
    tag = LETTERS[row_number()],
    sigma = paste0("sigma == ", sigma)
  )

moment_convergence <- moment_long |>
  filter(method != "truth") |>
  mutate(
    sigma = paste0("sigma == ", sigma),
    method = factor(method, levels = method_levels)
  ) |>
  ggplot(
    aes(
      x = n_classes,
      y = estimate,
      colour = method,
    )
  ) +
  geom_hline(
    aes(yintercept = true),
    data = moment_truth,
    linetype = 2,
    linewidth = 0.8
  ) +
  geom_text(
    aes(
      y = true,
      label = "true value"
    ),
    data = moment_truth,
    x = -Inf,
    hjust = -0.05,
    vjust = -0.6,
    size = 3,
    colour = "black",
    inherit.aes = FALSE
  ) +
  geom_vline(
    aes(xintercept = bins_1pct),
    data = moment_threshold,
    linetype = 3,
    linewidth = 0.8,
    colour = gq_colour
  ) +
  geom_text(
    aes(
      x = bins_1pct,
      label = paste0(bins_1pct, " bins, <1% error")
    ),
    data = moment_threshold,
    y = -Inf,
    hjust = -0.1,
    vjust = -0.8,
    size = 3,
    colour = gq_colour,
    inherit.aes = FALSE
  ) +
  xlab("number of bins") +
  geom_line(
    linewidth = 1
  ) +
  geom_text(
    aes(label = tag),
    data = panel_tags_moment,
    x = -Inf,
    y = Inf,
    hjust = 1.4,
    vjust = -1,
    size = 5,
    # fontface = "bold",
    colour = "black",
    inherit.aes = FALSE
  ) +
  coord_cartesian(clip = "off") +
  facet_wrap(
    moment ~ sigma,
    scales = "free_y",
    labeller = labeller(sigma = label_parsed)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

ggsave("figures/binning_moment_convergence.png",
       plot = moment_convergence,
       bg = "white",
       scale = 1,
       width = 9,
       height = 7)

```

![Convergence of each approximation to the mean (A--C) and standard deviation (D--F) of the activity-level distribution as the number of bins increases, for the quantile mean and Gaussian quadrature schemes, at three values of $\sigma$ (A, D: $\sigma = 0.5$; B, E: $\sigma = 1$; C, F: $\sigma = 2$). Horizontal dashed lines show the true values; vertical dotted lines mark the fewest bins at which the Gaussian quadrature estimate is within 1% of the truth.](../figures/binning_moment_convergence.png){#fig-binning-moments}

The quantile midpoint scheme underestimates the mean (since the midpoint value of each bin lies below its mean for this right-skewed distribution), and converges very slowly to the true estimate of the standard deviation as the number of bins increases. The quantile mean scheme approximates the mean well regardless of the number of bins (since the mean of the numerically integrated bin means is a very good approximator of the distribution mean), but likewise converges very slowly to the true variance. By contrast, the Gaussian quadrature scheme converges rapidly and consistently to both the mean and standard deviation at this task. Even with a very large sigma ($\sigma=2$ corresponds to a standard deviation of over a 50-fold difference in contact rates compared with log-mean), the Gaussian quadrature scheme is able to accurately approximate the standard deviation (to within 1% of the true value) with only 11 bins. By comparison, both evenly-sized-bin schemes significantly understimate the standard deviation (by a factor of more than 2) even with double the number of bins.

#### Approximating the eigenvalue

Finally, we asses the ability of each binning scheme to accurately estimate the eigenvalue of an activity-only contact matrix constructed from this distribution, again for the same three values of sigma. Here we display the true eigenvalue as a horizontal dashed line.

```{r}
plan(multisession, workers = 8)

eigenvalue_convergence <- expand_grid(
  bins = seq(2, 20, by = 1),
  sigma = c(0.5, 1, 2),
  method = c("quantile_midpoint",
             "quantile2_mean",
             "gaussquad")
) |>
  mutate(
    eigenvalue = furrr::future_pmap(
      list(bins, sigma, method),
      \(bins, sigma, method) map_to_eigen(n_classes = bins,
                                          sigma = sigma,
                                          method = method,
                                          assort = 0),
      .options = furrr_options(seed = TRUE)
    )
  ) |>
  # relabel methods for display (quantile2_mean is the deterministic
  # numerical-integration version of the quantile mean method)
  mutate(
    method = recode(
      method,
      quantile_midpoint = "quantile midpoint",
      quantile2_mean = "quantile mean",
      gaussquad = "Gaussian quadrature"
    )
  )
```

```{r}
# the true eigenvalue for the proportional mixing assumption, for each sigma
true_eigenval <- tibble(
  sigma = c(0.5, 1, 2)
) |>
  mutate(
    true = exp(1.5 * sigma ^ 2)
  )

eigenvalue_long <- eigenvalue_convergence |>
  unnest(
    eigenvalue
  ) |>
  left_join(
    true_eigenval,
    by = "sigma"
  )

# smallest number of bins for which the Gaussian quadrature estimate is
# within 1% of the true eigenvalue, for each sigma
convergence_threshold <- eigenvalue_long |>
  filter(
    method == "Gaussian quadrature",
    abs(eigenvalue - true) / true < 0.01
  ) |>
  group_by(sigma) |>
  summarise(
    bins_1pct = min(bins),
    .groups = "drop"
  ) |>
  mutate(
    sigma = paste0("sigma == ", sigma)
  )

# one panel letter per sigma facet (A-C)
panel_tags_eigen <- true_eigenval |>
  arrange(sigma) |>
  mutate(
    tag = LETTERS[row_number()],
    sigma = paste0("sigma == ", sigma)
  )

eigenvalue_convergence_plot <- eigenvalue_long |>
  mutate(
    sigma = paste0("sigma == ", sigma),
    method = factor(method, levels = method_levels)
  ) |>
  ggplot(
    aes(
      x = bins,
      y = eigenvalue,
      group = method,
      colour = method)
  ) +
  geom_hline(
    aes(yintercept = true),
    data = true_eigenval |> mutate(sigma = paste0("sigma == ", sigma)),
    linetype = 2,
    linewidth = 0.8
  ) +
  geom_text(
    aes(
      y = true,
      label = "true value"
    ),
    data = true_eigenval |> mutate(sigma = paste0("sigma == ", sigma)),
    x = -Inf,
    hjust = -0.05,
    vjust = -0.6,
    size = 3,
    colour = "black",
    inherit.aes = FALSE
  ) +
  geom_vline(
    aes(xintercept = bins_1pct),
    data = convergence_threshold,
    linetype = 3,
    linewidth = 0.8,
    colour = gq_colour
  ) +
  geom_text(
    aes(
      x = bins_1pct,
      label = paste0(bins_1pct, " bins, <1% error")
    ),
    data = convergence_threshold,
    y = -Inf,
    hjust = -0.1,
    vjust = -0.8,
    size = 3,
    colour = gq_colour,
    inherit.aes = FALSE
  ) +
  geom_line(
    linewidth = 1
  ) +
  geom_text(
    aes(label = tag),
    data = panel_tags_eigen,
    x = -Inf,
    y = Inf,
    hjust = 1.4,
    vjust = -1,
    size = 5,
    # fontface = "bold",
    colour = "black",
    inherit.aes = FALSE
  ) +
  coord_cartesian(clip = "off") +
  facet_wrap(
    ~ sigma,
    nrow = 1,
    scales = "free_y",
    labeller = labeller(sigma = label_parsed)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  xlab("number of bins") +
  ylab("dominant eigenvalue") +
  theme_minimal() +
  theme(legend.position = "bottom")

ggsave("figures/binning_eigenvalue_convergence.png",
       plot = eigenvalue_convergence_plot,
       bg = "white",
       scale = 1,
       width = 9,
       height = 4)
```

![Convergence to the truth of the dominant eigenvalue of the activity-only contact matrix as the number of bins increases, for the two binning schemes, at three values of $\sigma$ (A: $\sigma = 0.5$; B: $\sigma = 1$; C: $\sigma = 2$). Horizontal dashed lines show the true eigenvalue; vertical dotted lines mark the fewest bins at which the Gaussian quadrature estimate is within 1% of the truth.](../figures/binning_eigenvalue_convergence.png){#fig-binning-eigenvalue}

Across all values of $\sigma$ we see that the Gaussian quadrature scheme converges on the true eigenvalue of the activity-only contact matrix far more quickly than either evenly-sized-bin scheme (quantile midpoint and quantile mean). Even with the extreme $\sigma=2$, 12 bins are sufficient to approximate the eigenvalue to within 1%. With $\sigma=0.5$ (close to the estimate for the France data), 3 bins are sufficient to achieve the same high level of accuracy.

#### Convergence under assortative mixing

The eigenvalue convergence above uses a *proportionate* mixing activity matrix, for which an analytic eigenvalue is available. Our aim is to instead use *assortative* activity matrices, blending a proportionate and a fully-assortative matrix. These are governed by two further parameters: the width of the activity-assortativity kernel $\alpha$, and the overall degree of assortativity $\epsilon$. Here we confirm that the eigenvalue of the assortative activity matrix also converges as the number of bins increases, across a range of these parameters.

There is no analytic solution for the eigenvalue in the assortative case, so rather than comparing against a known truth we simply check that each scheme's estimate stops changing as the number of bins grows. As a reference we take the Gaussian quadrature estimate at the largest bin counts (the best available approximation to the truth), and mark the smallest number of bins from which the Gaussian quadrature estimate stays within 1% of it at every larger bin count.

The Gaussian quadrature estimate carries a small even/odd sawtooth as the number of bins increases: odd bin counts place a quadrature node exactly at the centre of the distribution while even counts do not, which slightly shifts the mass placed in the extreme upper tail (where the eigenvalue is most sensitive under a sharp assortativity kernel). Whether odd or even bins give a better approximation varies with parameter choices. To get a reliable reference despite this, we average the Gaussian quadrature estimates at the two largest bin counts (one odd, one even), which cancels most of the sawtooth. We also report the absolute percentage difference between those two evaluations, and flag any parameter combination where it exceeds 0.5% (indicating the sawtooth has not fully decayed by the maximum number of bins, so the reference is correspondingly less certain).

We fix $\sigma = 0.5$ (close to the France estimate) and sweep a $3 \times 3$ grid of kernel widths ($\alpha \in \{5, 10, 20\}$) and assortativity levels ($\epsilon \in \{0.5, 0.75, 1\}$). We exclude $\alpha = 0$ and $\epsilon = 0$, since either recovers the proportionate case already considered above. The matrices are constructed exactly as in analysis 3 (Tom's iterative-normalisation scheme for the fully-assortative matrix, blended with the proportionate matrix by $\epsilon$).

```{r}
#| message: false
#| warning: false

# fixed heterogeneity, and the assortativity grid (alpha = kernel width,
# epsilon = degree of assortativity). alpha = 0 / epsilon = 0 are excluded as
# they recover the proportionate case handled above.
sigma_assort <- 0.5
alpha_grid <- c(5, 10, 20)
epsilon_grid <- c(0.5, 0.75, 1)

assort_convergence <- expand_grid(
  bins = seq(2, 30, by = 1),
  alpha = alpha_grid,
  epsilon = epsilon_grid,
  method = c("quantile_midpoint",
             "quantile2_mean",
             "gaussquad")
) |>
  mutate(
    eigenvalue = furrr::future_pmap_dbl(
      list(bins, alpha, epsilon, method),
      \(bins, alpha, epsilon, method) activity_matrix_eigen(
        n_classes = bins,
        sigma = sigma_assort,
        alpha = alpha,
        epsilon = epsilon,
        method = method
      ),
      .options = furrr_options(seed = TRUE)
    )
  ) |>
  # relabel methods for display, matching the proportionate-case figures above
  mutate(
    method = recode(
      method,
      quantile_midpoint = "quantile midpoint",
      quantile2_mean = "quantile mean",
      gaussquad = "Gaussian quadrature"
    )
  )
```

```{r}
# facet labels (parsed expressions) for the alpha (rows) x epsilon (columns)
# grid. Ordered factors so the panels appear in ascending numeric order (rather
# than string-sorted, which would put alpha = 5 after alpha = 20/40).
assort_convergence <- assort_convergence |>
  mutate(
    alpha_lab = factor(
      paste0("alpha == ", alpha),
      levels = paste0("alpha == ", sort(alpha_grid))
    ),
    epsilon_lab = factor(
      paste0("epsilon == ", epsilon),
      levels = paste0("epsilon == ", sort(epsilon_grid))
    )
  )

# reference "converged" eigenvalue for each (alpha, epsilon). There is no
# analytic truth in the assortative case, so we approximate it from the largest
# bin counts. The Gaussian quadrature estimate carries a small even/odd
# Gauss-Hermite sawtooth (odd bin counts place a node at the distribution centre,
# even counts do not), so a single bin count is a slightly biased reference; we
# average the two largest bin counts (one odd, one even) to cancel most of it.
ref_bins <- sort(unique(assort_convergence$bins), decreasing = TRUE)[1:2]

assort_reference <- assort_convergence |>
  filter(method == "Gaussian quadrature", bins %in% ref_bins) |>
  group_by(alpha, epsilon, alpha_lab, epsilon_lab) |>
  summarise(reference = mean(eigenvalue), .groups = "drop")

# residual uncertainty in that reference: the absolute % difference between the
# two Gaussian quadrature evaluations we average (the largest two bin counts).
# Where this is small the reference is reliable; where the sawtooth persists
# (strong assortativity) it quantifies how much the two evaluations still differ.
assort_ref_diff <- assort_convergence |>
  filter(method == "Gaussian quadrature", bins %in% ref_bins) |>
  left_join(assort_reference |> select(alpha, epsilon, reference),
            by = c("alpha", "epsilon")) |>
  group_by(alpha, epsilon) |>
  summarise(
    eig_lower_bin = eigenvalue[bins == min(ref_bins)],
    eig_upper_bin = eigenvalue[bins == max(ref_bins)],
    pct_diff = 100 * abs(eig_upper_bin - eig_lower_bin) / first(reference),
    within_0.5pct = pct_diff < 0.5,
    .groups = "drop"
  )

# report the per-panel differences, and flag any where the two evaluations we
# average still disagree by more than 0.5% (i.e. the reference is less certain)
assort_ref_diff

if (any(!assort_ref_diff$within_0.5pct)) {
  warning(
    sprintf(
      "reference from bins %d/%d differs by >0.5%% for: %s (residual sawtooth)",
      ref_bins[2], ref_bins[1],
      paste(
        sprintf("alpha=%g,epsilon=%g (%.2f%%)",
                assort_ref_diff$alpha[!assort_ref_diff$within_0.5pct],
                assort_ref_diff$epsilon[!assort_ref_diff$within_0.5pct],
                assort_ref_diff$pct_diff[!assort_ref_diff$within_0.5pct]),
        collapse = "; "
      )
    )
  )
}

# number of bins from which the Gaussian quadrature estimate stays within 1% of
# the reference, for each (alpha, epsilon). The estimate carries an even/odd
# sawtooth in the assortative case, so the *first* bin count to fall within 1% is
# a poor guide to how many bins are actually needed: adding one more bin can push
# the error back above 1%. We instead take one more than the *largest* bin count
# whose error still exceeds 1% -- the smallest bin count from which the estimate
# stays within tolerance for good, robust to the sawtooth crossing back out. If no
# bin count exceeds 1% (converged from the first bin), we take the smallest count.
min_assort_bins <- min(assort_convergence$bins)

assort_threshold <- assort_convergence |>
  filter(method == "Gaussian quadrature") |>
  left_join(
    assort_reference |> select(alpha, epsilon, reference),
    by = c("alpha", "epsilon")
  ) |>
  mutate(over_1pct = abs(eigenvalue - reference) / reference > 0.01) |>
  group_by(alpha, epsilon, alpha_lab, epsilon_lab) |>
  summarise(
    bins_1pct = if (any(over_1pct)) max(bins[over_1pct]) + 1L else min_assort_bins,
    .groups = "drop"
  )

assort_convergence_plot <- assort_convergence |>
  mutate(method = factor(method, levels = method_levels)) |>
  ggplot(
    aes(
      x = bins,
      y = eigenvalue,
      group = method,
      colour = method
    )
  ) +
  geom_hline(
    aes(yintercept = reference),
    data = assort_reference,
    linetype = 2,
    linewidth = 0.8
  ) +
  geom_text(
    aes(
      y = reference,
      label = "reference (mean of 29 & 30 bins)"
    ),
    data = assort_reference,
    x = -Inf,
    hjust = -0.05,
    vjust = -0.6,
    size = 2.7,
    colour = "black",
    inherit.aes = FALSE
  ) +
  geom_vline(
    aes(xintercept = bins_1pct),
    data = assort_threshold,
    linetype = 3,
    linewidth = 0.8,
    colour = gq_colour
  ) +
  geom_text(
    aes(
      x = bins_1pct,
      label = paste0("<1% from ", bins_1pct, " bins")
    ),
    data = assort_threshold,
    y = -Inf,
    hjust = -0.1,
    vjust = -0.8,
    size = 2.7,
    colour = gq_colour,
    inherit.aes = FALSE
  ) +
  geom_line(
    linewidth = 1
  ) +
  coord_cartesian(clip = "off") +
  facet_wrap(
    alpha_lab ~ epsilon_lab,
    scales = "free_y",
    labeller = label_parsed
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  xlab("number of bins") +
  ylab("dominant eigenvalue") +
  theme_minimal() +
  theme(legend.position = "bottom")

ggsave("figures/binning_eigenvalue_convergence_assortative.png",
       plot = assort_convergence_plot,
       bg = "white",
       scale = 1,
       width = 10,
       height = 9)
```

![Convergence of the dominant eigenvalue of the *assortative* activity matrix as the number of bins increases, for the three binning schemes, at fixed $\sigma = 0.5$. Panels vary the activity-assortativity kernel width $\alpha$ (rows: 5, 10, 20) and the degree of assortativity $\epsilon$ (columns: 0.5, 0.75, 1). As there is no analytic truth in the assortative case, the horizontal dashed line shows the reference eigenvalue --- the mean of the Gaussian quadrature estimates at the two largest bin counts (29 and 30), which cancels most of the even/odd sawtooth --- and the vertical dotted line marks the smallest number of bins from which the Gaussian quadrature estimate stays within 1% of it at every larger bin count.](../figures/binning_eigenvalue_convergence_assortative.png){#fig-binning-eigenvalue-assortative}

As in the proportionate case, the Gaussian quadrature scheme settles far more quickly than either evenly-sized-bin scheme, across the whole grid of assortativity parameters. The two quantile schemes converge slowly and, at strong assortativity (large $\alpha$ and $\epsilon$), are still visibly biased even at 30 bins. Convergence is fastest at weak assortativity and slows as both $\alpha$ and $\epsilon$ increase --- in the most strongly-assortative corner ($\alpha = 20$, $\epsilon = 1$) the eigenvalue is most sensitive to the placement of bins in the extreme tail of the activity distribution, and this is also where the even/odd sawtooth persists most stubbornly: the two evaluations we average for the reference still differ by around 0.07% at 30 bins, compared with under 0.01% for the milder combinations ($\alpha \le 10$). Nonetheless, once the sawtooth is averaged out the Gaussian quadrature estimate stops changing appreciably within a modest number of bins across the whole grid: even in the most strongly-assortative corner ($\alpha = 20$, $\epsilon = 1$) 10 bins are enough for the estimate to stay within 1% of the reference, comfortably below the number used in the real-world analysis.
