---
title: "Dominant eigenvalue vs activity heterogeneity"
format: html
editor: visual
---

# Dominant eigenvalue of the activity-structured contact matrix

This document remakes the MATLAB figure produced by `matlab/contact_models.m` (`figure(4)`): the dominant eigenvalue $\lambda$ of an assortative, activity-structured contact matrix as a function of the activity heterogeneity $\sigma$ (the standard deviation of the log activity-level distribution), with a separate line for each value of the assortativity kernel width $\alpha$.

Rather than the MATLAB discretisation (100 evenly-sized quantile bins), it uses the R functions from the analysis 3 pipeline: the activity classes are defined by Gauss-Hermite quadrature (`gaussquad_classes()`), the matrix is built with Tom's iterative normalisation method (`make_toms_contact_matrix()`, a port of `matlab/makeContactMatrix_AM_Tom.m`), and the dominant eigenvalue is extracted with `get_eigenval()`.

The second half of the document confirms that the quadrature scheme is a more accurate approximation of the continuous activity model than the even-bin scheme, by comparing both against the analytic answer available in the proportionate-mixing ($\alpha = 0$) case.

```{r}
#| message: false
#| warning: false
source("R/packages.R")
source("R/functions.R")

# seed for the Monte Carlo binning method used in the comparison below
set.seed(2026-07-08)
```

## The figure

We sweep the dominant eigenvalue over a grid of activity heterogeneity $\sigma$ and assortativity kernel width $\alpha$. The $\alpha$ values match those plotted in the MATLAB figure (`Alpha_arr = 0:2:40`, every 5th value). We use 100 quadrature points, matching the MATLAB bin count; this gives visually converged curves even for the sharpest ($\alpha = 40$) kernel.

```{r}
# activity heterogeneity (sd of the log activity distribution)
sigma_grid <- seq(0, 1, by = 0.025)

# assortativity kernel width, matching the values plotted in the MATLAB figure
alpha_values <- c(0, 10, 20, 30, 40)

# number of quadrature points used to approximate the continuous activity
# distribution
n_classes <- 100
```

For each $(\sigma, \alpha)$ we build the activity classes by quadrature, construct the assortative matrix, and take its dominant eigenvalue. When $\sigma = 0$ everyone shares the same activity level, so the matrix is homogeneous with dominant eigenvalue 1 for any $\alpha$.

```{r}
dominant_eigenvalue <- function(sigma, alpha, n_classes) {
  if (sigma == 0) {
    return(1)
  }
  activity_bins <- gaussquad_classes(sigma = sigma, n_classes = n_classes)
  M <- make_toms_contact_matrix(
    v = activity_bins$activity,
    p_pop = activity_bins$fraction,
    sigma = sigma,
    alpha = alpha
  )
  get_eigenval(M)
}

eigen_grid <- expand_grid(
  sigma = sigma_grid,
  alpha = alpha_values
) |>
  mutate(
    lambda = pmap_dbl(
      list(sigma, alpha),
      dominant_eigenvalue,
      n_classes = n_classes
    ),
    alpha = factor(alpha, levels = alpha_values)
  )
```

```{r}
eigenvalue_plot <- eigen_grid |>
  ggplot(aes(x = sigma, y = lambda, colour = alpha)) +
  geom_line(linewidth = 0.8) +
  scale_colour_viridis_d(end = 0.9, direction = -1) +
  labs(
    x = expression("activity heterogeneity (" * sigma * ")"),
    y = expression("dominant eigenvalue (" * lambda * ")"),
    colour = expression("assortativity (" * alpha * ")")
  ) +
  theme_minimal()

eigenvalue_plot

ggsave("figures/eigenvalue_vs_sigma.png",
       plot = eigenvalue_plot,
       bg = "white",
       width = 7,
       height = 5,
       dpi = 300)
```

## Confirming the quadrature is more accurate

The curves above sit higher than those in the MATLAB figure. This is expected if the Gauss-Hermite quadrature is the more accurate scheme, because the MATLAB even-bin discretisation truncates the heavy upper tail of the lognormal activity distribution --- and it is exactly that tail that inflates $E[v^2]$ and hence the dominant eigenvalue.

To confirm this we recompute the curves under two activity discretisations with the same number of bins:

-   **Gauss-Hermite quadrature** (`gaussquad_classes()`) --- unequal weights at quadrature nodes; the accurate method used in analysis 3.
-   **even bins, MATLAB** --- the inverse-CDF value at each bin's midpoint quantile with uniform weights, a faithful port of the MATLAB discretisation (`v = logninv(x, 0, sigma)` at bin midpoints $x$).

The $\alpha = 0$ (proportionate mixing) case has a known analytic dominant eigenvalue, $\exp(\sigma^2)$, giving an exact reference to judge accuracy against.

```{r}
# activity classes as an (activity, fraction) tibble for a given method and sigma
activity_classes <- function(method, sigma) {
  switch(
    method,
    # accurate quadrature scheme used in analysis 3
    gaussquad = gaussquad_classes(sigma = sigma, n_classes = n_classes),
    # faithful port of the MATLAB discretisation: inverse-CDF at midpoint
    # quantiles, uniform population weights
    matlab = tibble(
      activity = qlnorm((seq_len(n_classes) - 0.5) / n_classes, 0, sigma),
      fraction = 1 / n_classes
    )
  )
}

dominant_eigenvalue_method <- function(method, sigma, alpha) {
  if (sigma == 0) {
    return(1)
  }
  bins <- activity_classes(method, sigma)
  M <- make_toms_contact_matrix(
    v = bins$activity,
    p_pop = bins$fraction,
    sigma = sigma,
    alpha = alpha
  )
  get_eigenval(M)
}

method_labels <- c(
  gaussquad = "Gauss-Hermite quadrature",
  matlab    = "even bins, MATLAB"
)

eigen_grid_methods <- expand_grid(
  method = names(method_labels),
  sigma  = sigma_grid,
  alpha  = alpha_values
) |>
  mutate(
    lambda = pmap_dbl(list(method, sigma, alpha), dominant_eigenvalue_method),
    method = factor(method_labels[method], levels = method_labels),
    alpha  = factor(alpha, levels = alpha_values)
  )
```

```{r}
comparison_plot <- eigen_grid_methods |>
  ggplot(aes(x = sigma, y = lambda, colour = alpha, linetype = method)) +
  geom_line(linewidth = 0.7) +
  scale_colour_viridis_d(end = 0.9, direction = -1) +
  labs(
    x = expression("activity heterogeneity (" * sigma * ")"),
    y = expression("dominant eigenvalue (" * lambda * ")"),
    colour = expression("assortativity (" * alpha * ")"),
    linetype = "binning method"
  ) +
  theme_minimal()

comparison_plot

ggsave("figures/eigenvalue_vs_sigma_binning_comparison.png",
       plot = comparison_plot,
       bg = "white",
       width = 8,
       height = 5,
       dpi = 300)
```

### Numerical accuracy check

For $\alpha = 0$ the exact dominant eigenvalue is $\exp(\sigma^2)$; at $\sigma = 1$ that is $\exp(1) = 2.71828$. The quadrature scheme recovers it essentially exactly, while the MATLAB even-bin scheme falls short by around 12% (and the MC bin-mean method, which captures more of the tail than a single midpoint value, by around 6%). The "even bins, MATLAB" curve also reproduces the saved MATLAB figure (`matlab/figures/eigenvalue.png`), confirming that the shift between the two R figures is due entirely to the binning scheme.

```{r}
accuracy_check <- eigen_grid_methods |>
  filter(sigma == 1) |>
  select(method, alpha, lambda) |>
  pivot_wider(names_from = method, values_from = lambda)

knitr::kable(accuracy_check, digits = 4)
```

The analytic reference at $\alpha = 0$, $\sigma = 1$ is $\exp(1) = `r sprintf("%.4f", exp(1))`$.
