---
title: "Final size vs activity heterogeneity"
format: html
editor: visual
---

# Final size of the activity-structured model

This is the final-size counterpart of `R/eigenvalue_vs_sigma.qmd`: the same sweep over activity heterogeneity $\sigma$ and assortativity kernel width $\alpha$, but plotting the overall final size of the epidemic rather than the relative $R_0$.

Heterogenising the population always increases $R_0$ (the eigenvalue figure), but its effect on the final size pulls in two directions. Heterogeneity concentrates infection in the most active individuals, which *lowers* the attack rate at a given $R_0$; it also *raises* $R_0$, which pushes the attack rate up. Which one wins depends on what is being held fixed, so the two panels hold different things fixed and answer different questions:

-   **Panel A --- homogeneous $R_0$ held at 0.9.** The transmission probability $\beta$ is fixed, at a value that could not sustain an epidemic in a homogeneous population. As $\sigma$ rises the realised $R_0$ climbs through 1 and the final size lifts off zero. Assortativity brings that threshold forward, because it raises $R_0$ faster.
-   **Panel B --- realised $R_0$ held at 3.** Here $\beta$ is *lowered* as $\sigma$ rises, exactly enough to keep the realised $R_0$ at 3. With the $R_0$ effect held out, only the concentrating effect remains, and the final size falls with $\sigma$ --- further for more assortative mixing.

At $\sigma = 0$ there is no variation in activity level for the assortativity kernel to sort on, so every $\alpha$ gives the same final size and the curves in each panel start from a common point.

The shallow strip under each panel makes the difference explicit: it shows both $R_0$s over the same $\sigma$ axis, with the one held constant drawn flat in grey and the one left free coloured by $\alpha$. The dotted line marks the epidemic threshold $R_0 = 1$.

```{r}
#| message: false
#| warning: false
source("R/packages.R")
source("R/functions.R")
```

## Setup

The $\sigma$ and $\alpha$ grids, the number of quadrature points, and the line colours all match the eigenvalue figure, so the two can be read side by side.

```{r}
# activity heterogeneity (sd of the log activity distribution)
sigma_grid <- seq(0, 1, by = 0.025)

# assortativity kernel width, matching the eigenvalue figure
alpha_values <- c(0, 10, 20, 30, 40)

# number of quadrature classes used to discretise the activity distribution,
# matching the eigenvalue figure. This many is needed in the demanding corner of
# the sweep: at sigma = 1 with alpha = 40, 20 classes is 2.4% short of the
# converged final size, while 100 agrees with it to 5 decimal places
n_classes <- 100

# fully-assortative blending, matching the main analysis
epsilon <- 1

# the R0 each panel holds fixed: panel A fixes the homogeneous R0 below the
# epidemic threshold, panel B fixes the realised R0 well above it
r0_homogeneous <- 0.9
r0_realised <- 3

# colours for the alpha lines: pink (low alpha) through to dark purple (high
# alpha), using the darker portion of the ColorBrewer RdPu palette
alpha_colours <- colorRampPalette(
  RColorBrewer::brewer.pal(9, "RdPu")[4:9]
)(length(alpha_values))

# the R0 held constant in each panel is drawn once in grey (it is the same for
# every alpha); the R0 left free is coloured by alpha, and line type says which
# of the two R0s a line is
r0_linetypes <- c("Realised" = "solid", "Homogeneous" = "dashed")
constant_colour <- "grey40"
```

The activity matrix has a dominant eigenvalue of 1 in the homogeneous case, so the transmission probability $\beta$ *is* the homogeneous $R_0$, and the realised $R_0$ is $\beta$ times the relative $R_0$ $\lambda(\sigma, \alpha)$ plotted in the eigenvalue figure. That makes both panels a matter of choosing $\beta$: fix it at 0.9 for panel A, or set it to $3 / \lambda$ for panel B.

```{r}
# relative R0 for each (sigma, alpha): the quantity plotted in the eigenvalue
# figure, and the bridge between the two R0s
lambda_grid <- expand_grid(
  sigma = sigma_grid,
  alpha = alpha_values
) |>
  mutate(
    lambda = pmap_dbl(
      list(sigma, alpha),
      \(sigma, alpha) get_eigenval(
        make_activity_matrix(
          n_activity_bins = n_classes,
          sigma = sigma,
          alpha = alpha,
          epsilon = epsilon
        )
      )
    )
  )
```

Final sizes are computed with `activity_final_size()` (defined in `R/functions.R`) --- the activity-only counterpart of the `age_activity_final_size()` used in analysis 3, so they are computed exactly as they are there.

```{r}
add_final_size <- function(data) {
  data |>
    mutate(
      final_size = pmap_dbl(
        list(sigma, alpha, beta),
        \(sigma, alpha, beta) activity_final_size(
          sigma = sigma,
          alpha = alpha,
          epsilon = epsilon,
          beta = beta,
          n_activity_bins = n_classes
        )$final_size
      ),
      alpha = factor(alpha, levels = alpha_values)
    )
}

# panel A: beta fixed, so the realised R0 rises with sigma
panel_a_data <- lambda_grid |>
  mutate(
    beta = r0_homogeneous,
    realised = beta * lambda
  ) |>
  add_final_size()

# panel B: beta lowered as sigma rises, holding the realised R0 at its target
panel_b_data <- lambda_grid |>
  mutate(
    beta = r0_realised / lambda,
    realised = r0_realised
  ) |>
  add_final_size()
```

## The figure

```{r}
# the final size panel: the top three quarters of each column. Its x axis is
# suppressed, because it shares the R0 strip's axis directly below it
final_size_panel <- function(data, title, tag) {
  data |>
    ggplot(aes(x = sigma, y = final_size, colour = alpha)) +
    geom_line(linewidth = 0.8) +
    scale_colour_manual(values = alpha_colours) +
    scale_y_continuous(labels = scales::label_percent()) +
    labs(
      y = "Overall final size",
      colour = expression("Assortativity (" * alpha * ")"),
      title = title,
      tag = tag
    ) +
    theme_minimal() +
    theme(
      axis.title.x = element_blank(),
      axis.text.x = element_blank()
    )
}

# the R0 strip below it: `varying` is the R0 this panel lets move (coloured by
# alpha), `constant` the one it holds fixed (one grey line, not one per alpha)
r0_panel <- function(varying, constant, varying_type, constant_type,
                     show_legend = FALSE) {
  ggplot() +
    geom_hline(yintercept = 1, linetype = "dotted", colour = "grey70") +
    geom_line(
      data = varying,
      aes(x = sigma, y = r0, colour = alpha, linetype = varying_type),
      linewidth = 0.7
    ) +
    geom_line(
      data = constant,
      aes(x = sigma, y = r0, linetype = constant_type),
      colour = constant_colour,
      linewidth = 0.7
    ) +
    scale_colour_manual(values = alpha_colours, guide = "none") +
    scale_linetype_manual(
      values = r0_linetypes,
      name = NULL,
      limits = names(r0_linetypes)
    ) +
    # the two strips would otherwise contribute two identical line type
    # legends, which patchwork does not merge, so only one of them draws it
    guides(linetype = if (show_legend) {
      guide_legend(
        override.aes = list(colour = constant_colour, linewidth = 0.7)
      )
    } else {
      "none"
    }) +
    labs(
      x = expression("Activity heterogeneity (" * sigma * ")"),
      y = expression(R[0])
    ) +
    theme_minimal()
}

column_a <- final_size_panel(
  panel_a_data,
  bquote("Homogeneous " * R[0] * " held at " * .(r0_homogeneous)),
  "A"
) /
  r0_panel(
    varying = mutate(panel_a_data, r0 = realised),
    constant = tibble(sigma = sigma_grid, r0 = r0_homogeneous),
    varying_type = "Realised",
    constant_type = "Homogeneous",
    show_legend = TRUE
  ) +
  plot_layout(heights = c(3, 1))

column_b <- final_size_panel(
  panel_b_data,
  bquote("Realised " * R[0] * " held at " * .(r0_realised)),
  "B"
) /
  r0_panel(
    varying = mutate(panel_b_data, r0 = beta),
    constant = tibble(sigma = sigma_grid, r0 = r0_realised),
    varying_type = "Homogeneous",
    constant_type = "Realised"
  ) +
  plot_layout(heights = c(3, 1))

final_size_plot <- (column_a | column_b) +
  plot_layout(guides = "collect")

final_size_plot

ggsave("figures/final_size_vs_sigma.png",
       plot = final_size_plot,
       bg = "white",
       width = 11,
       height = 6,
       dpi = 300)
```

The numbers behind the curves, at four values of $\sigma$.

```{r}
bind_rows(
  `Homogeneous R0 = 0.9` = panel_a_data,
  `Realised R0 = 3` = panel_b_data,
  .id = "Panel"
) |>
  filter(sigma %in% c(0.25, 0.5, 0.75, 1)) |>
  transmute(
    Panel,
    sigma,
    alpha,
    `Homogeneous R0` = round(beta, 3),
    `Realised R0` = round(realised, 3),
    `Final size` = round(final_size, 3)
  ) |>
  knitr::kable()
```
