---
title: "Final size vs activity heterogeneity"
format: html
editor: visual
---

# Final size of the activity-structured model

This is the final-size counterpart of `R/eigenvalue_vs_sigma.qmd`: the same sweep over activity heterogeneity $\sigma$ and assortativity kernel width $\alpha$, but plotting the overall final size of the epidemic rather than the relative $R_0$.

Heterogenising the population always increases $R_0$ (the eigenvalue figure), but its effect on the final size pulls in two directions. Heterogeneity concentrates infection in the most active individuals, which *lowers* the attack rate at a given $R_0$; it also *raises* $R_0$, which pushes the attack rate up. Which one wins depends on what is being held fixed, so we make two figures that hold different things fixed and answer different questions. In each, panel A shows the $R_0$ that is left free to vary and panel B the overall final size:

-   **Homogeneous $R_0$ held at 0.9 (main text).** The transmission probability $\beta$ is fixed, at a value that could not sustain an epidemic in a homogeneous population. As $\sigma$ rises the realised $R_0$ climbs through 1 and the final size lifts off zero. Assortativity brings that threshold forward, because it raises $R_0$ faster. At larger $\sigma$ the concentrating effect catches up with the rise in $R_0$ and the final size turns back down, so we extend $\sigma$ to 1.5 to show this.
-   **Realised $R_0$ held at 3 (supplement).** Here $\beta$ is *lowered* as $\sigma$ rises, exactly enough to keep the realised $R_0$ at 3. With the $R_0$ effect held out, only the concentrating effect remains, and the final size falls with $\sigma$ --- further for more assortative mixing.

At $\sigma = 0$ there is no variation in activity level for the assortativity kernel to sort on, so every $\alpha$ gives the same final size and the curves in each panel of each figure start from a common point.

Panel A, above the final size, makes the difference explicit: it shows the $R_0$ left free to vary, coloured by $\alpha$, over the same $\sigma$ axis (the realised $R_0$ in the first figure, the homogeneous $R_0$ in the second). The dotted line marks the epidemic threshold $R_0 = 1$. Note that panel A of the first figure shows the absolute realised $R_0$, not the relative $R_0$ plotted in the eigenvalue figure. The $R_0$ held fixed in each figure is not drawn; it is given in the caption.

```{r}
#| message: false
#| warning: false
source("R/packages.R")
source("R/functions.R")
```

## Setup

The $\alpha$ grid, the number of quadrature points, and the line colours all match the eigenvalue figure, so the two can be read side by side. The $\sigma$ grid extends further, to show the final size turning down.

```{r}
# activity heterogeneity (sd of the log activity distribution), extending past
# the eigenvalue figure's range so the final size can be seen to peak and fall
sigma_grid <- seq(0, 1.5, by = 0.025)

# assortativity kernel width, matching the eigenvalue figure
alpha_values <- c(0, 10, 20, 30, 40)

# number of quadrature classes used to discretise the activity distribution,
# matching the eigenvalue figure. This many is needed in the demanding corner of
# the sweep: at sigma = 1 with alpha = 40, 20 classes is 2.4% short of the
# converged final size, while 100 agrees with it to 5 decimal places
n_classes <- 100

# fully-assortative blending, matching the main analysis
epsilon <- 1

# the R0 each figure holds fixed: the first fixes the homogeneous R0 below the
# epidemic threshold, the second fixes the realised R0 well above it
r0_homogeneous <- 0.9
r0_realised <- 3

# colours for the alpha lines: pink (low alpha) through to dark purple (high
# alpha), using the darker portion of the ColorBrewer RdPu palette
alpha_colours <- colorRampPalette(
  RColorBrewer::brewer.pal(9, "RdPu")[4:9]
)(length(alpha_values))
```

The activity matrix has a dominant eigenvalue of 1 in the homogeneous case, so the transmission probability $\beta$ *is* the homogeneous $R_0$, and the realised $R_0$ is $\beta$ times the relative $R_0$ $\lambda(\sigma, \alpha)$ plotted in the eigenvalue figure. That makes both figures a matter of choosing $\beta$: fix it at 0.9 for the first, or set it to $3 / \lambda$ for the second.

```{r}
# relative R0 for each (sigma, alpha): the quantity plotted in the eigenvalue
# figure, and the bridge between the two R0s
lambda_grid <- expand_grid(
  sigma = sigma_grid,
  alpha = alpha_values
) |>
  mutate(
    lambda = pmap_dbl(
      list(sigma, alpha),
      \(sigma, alpha) get_eigenval(
        make_activity_matrix(
          n_activity_bins = n_classes,
          sigma = sigma,
          alpha = alpha,
          epsilon = epsilon
        )
      )
    )
  )
```

Final sizes are computed with `activity_final_size()` (defined in `R/functions.R`) --- the activity-only counterpart of the `age_activity_final_size()` used in analysis 3, so they are computed exactly as they are there.

```{r}
add_final_size <- function(data) {
  data |>
    mutate(
      final_size = pmap_dbl(
        list(sigma, alpha, beta),
        \(sigma, alpha, beta) activity_final_size(
          sigma = sigma,
          alpha = alpha,
          epsilon = epsilon,
          beta = beta,
          n_activity_bins = n_classes
        )$final_size
      ),
      alpha = factor(alpha, levels = alpha_values)
    )
}

# homogeneous R0 held constant: beta fixed, so the realised R0 rises with sigma
homogeneous_data <- lambda_grid |>
  mutate(
    beta = r0_homogeneous,
    realised = beta * lambda
  ) |>
  add_final_size()

# realised R0 held constant: beta lowered as sigma rises, holding the realised
# R0 at its target
realised_data <- lambda_grid |>
  mutate(
    beta = r0_realised / lambda,
    realised = r0_realised
  ) |>
  add_final_size()
```

## The figures

```{r}
# panel A, the R0 this figure lets move, one line per alpha: the top two fifths
# of each figure. Its x axis is suppressed, because it shares the final size
# panel's axis directly below it. The R0 it holds fixed is given in the caption
# rather than drawn
r0_panel <- function(data, y_label) {
  data |>
    ggplot(aes(x = sigma, y = r0, colour = alpha)) +
    geom_hline(yintercept = 1, linetype = "dotted", colour = "grey70") +
    geom_line(linewidth = 0.8) +
    scale_colour_manual(values = alpha_colours, guide = "none") +
    expand_limits(y = 0) +
    labs(y = y_label) +
    theme_minimal() +
    theme(
      axis.title.x = element_blank(),
      axis.text.x = element_blank()
    )
}

# panel B, the final size, below it. `y_limits` optionally fixes the y axis
# range (NULL lets ggplot choose)
final_size_panel <- function(data, y_limits = NULL) {
  data |>
    ggplot(aes(x = sigma, y = final_size, colour = alpha)) +
    geom_line(linewidth = 0.8) +
    scale_colour_manual(values = alpha_colours) +
    scale_y_continuous(labels = scales::label_percent(), limits = y_limits) +
    labs(
      x = expression("Activity heterogeneity (" * sigma * ")"),
      y = "Overall final size",
      colour = expression("Assortativity (" * alpha * ")")
    ) +
    theme_minimal()
}

# homogeneous R0 held constant, with the realised R0 free (main text)
homogeneous_plot <- r0_panel(
  mutate(homogeneous_data, r0 = realised),
  y_label = expression(R[0])
) /
  final_size_panel(homogeneous_data) +
  plot_layout(heights = c(2, 3), guides = "collect") +
  plot_annotation(tag_levels = "A")

homogeneous_plot

ggsave("figures/final_size_vs_sigma_homogeneous.png",
       plot = homogeneous_plot,
       bg = "white",
       width = 7,
       height = 7.5,
       dpi = 300)

# realised R0 held constant, with the homogeneous R0 free (supplement). The
# final size spans most of the range here, so show the full 0-100% axis
realised_plot <- r0_panel(
  mutate(realised_data, r0 = beta),
  y_label = expression("Homogeneous " * R[0])
) /
  final_size_panel(realised_data, y_limits = c(0, 1)) +
  plot_layout(heights = c(2, 3), guides = "collect") +
  plot_annotation(tag_levels = "A")

realised_plot

ggsave("figures/final_size_vs_sigma_realised.png",
       plot = realised_plot,
       bg = "white",
       width = 7,
       height = 7.5,
       dpi = 300)
```

The numbers behind the curves, at a selection of values of $\sigma$.

```{r}
bind_rows(
  `Homogeneous R0 = 0.9` = homogeneous_data,
  `Realised R0 = 3` = realised_data,
  .id = "Figure"
) |>
  filter(sigma %in% c(0.25, 0.5, 1, 1.5)) |>
  transmute(
    Figure,
    sigma,
    alpha,
    `Homogeneous R0` = round(beta, 3),
    `Realised R0` = round(realised, 3),
    `Final size` = round(final_size, 3)
  ) |>
  knitr::kable()
```
